### 2022-01-09 manuscript.Rmd
Add manuscript.Rmd

---
title: Template for Oxford University Press papers
author:
  - name: Alpha Beta
    email: alpha@university.ac.uk
    corresponding_author: TRUE # Automatic footnote using corresponding_author
    affiliation:
      - ABC
      - DEF
    biography: |
      A researcher.
      #    biography_pic: "/PATH/TO/image.png"
      #    biography_pic_width: "1in"
  - name: Beta Gamma
    email: beta@university.ac.uk
    affiliation: DEF
    footnote: 1 # Footnote using reference
  - name: Zeta Epsilon
    email: zeta@institution.ac.uk
    affiliation: XYZ
    footnote:
        - 1
        - 2
address:
  - code: ABC
    department: Department of Biology
    organization: University
  - code: DEF
    address: University
  - code: XYZ
    address: Institution
footnote:
  - code: 1
    text: "Equal contribution"
  - code: 2
    text: "Current email address: [alpha@university.ac.uk](mailto:alpha@university.ac.uk)"
abstract:
  - "**Objective:**  To Investigate, and identify, the factors that may influence patient survival of coronavirus disease 19."
  - "More"
  - "**Motivation:** You can also have some paragraphs start with bold face."
keywords:
  - mathematics
  - survival
  - prognostics
competing_interests: |
  There are no competing interest.
author_contributions: |
  Equal contributions
acknowledgements: |
  This is an acknowledgement.

  It consists of two paragraphs.

bibliography: manuscript/mybibfile.bib # Includes refs in OUP example template
# When using `citation_package = "default"`, a CSL file can be used
# csl: https://www.zotero.org/styles/oxford-university-press-note
output:
  rticles::oup_article:
    oup_version: 1 # 1 = 2020 CTAN OUP CLS package; 0 = 2009 OUP CLS package
    keep_tex: FALSE
    #journal: "Journal Name"
    document_style: "contemporary" # Can be contemporary, modern, traditional
    papersize: "large" # Can be large, medium, small
    #citation_package: "default" # Uncomment when using a CSL; default "natbib"
    namedate: TRUE # Set FALSE to use numeric refs; Default FALSE
    #number_sections: FALSE # Uncomment to not number sections; default TRUE
    #number_lines: TRUE # Use `lineno` package - Default FALSE
    #number_lines_options: ["mathlines","switch"]  # Options for latex lineno package.
    #onecolumn: TRUE # Uncomment for one column format; default FALSE
    extra_dependencies:
      - booktabs # to use with knitr::kable() example below

# Example of pandoc's variable
# urlcolor: orange
# linkcolor: green
# citecolor: red
header-includes:
  - \theoremstyle{thmstyleone}
  - \newtheorem{theorem}{Theorem}
  - \newtheorem{proposition}[theorem]{Proposition}
  - \theoremstyle{thmstyletwo}
  - \newtheorem{example}{Example}
  - \newtheorem{remark}{Remark}
  - \theoremstyle{thmstylethree}
  - \newtheorem{definition}{Definition}
---

<!--- Global Settings
    th -> top or here
--->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.pos = 'th',
  out.width = '100%', dpi = 300, fig.env="figure")

options(knitr.table.format = "latex") # For kable tables to write LaTeX table directly
```

<!--- imports --->
```{r include = FALSE}
library(xtable)
library(dplyr)
library(magrittr)
```

<!--- external functios --->
```{r}
sys.source(file = 'R/functions/StudyData.R', envir = knitr::knit_global())
```

# Introduction

This template is based on the generic OUP authoring template available on CTAN under [oup-authoring-template](https://www.ctan.org/pkg/oup-authoring-template). The CTAN template includes LaTeX documentation and a sample LaTeX document that provide far more details regarding the full functionality of the format. Here, only the basic functioning of the Rmarkdown adaptation of the format is demonstrated.

## A subsection

A numbered list:

1) First point
2) Second point
    - Subpoint

A bullet list:

* First point
* Second point

## Notes

* Extra white space in document will tend to disappear as text is filled in.
* Code blocks tend to generate lots of empty white space when `echo=TRUE` for some reason.

# Methods

## Data Source
The data is a derivative of the ... It is a cohort study wherein ...

## Participants

* Study setting
* Eligibility criteria

## Outcome

* Died

## Predictors

# Literature citations

By default, citations are handled by `natbib` using a numeric citation format. To use name-date citations, sets `namedate: TRUE` in the YAML header.

Here are two sample references:

* **author (year) example:** @horvath2018dna showed some really cool things. Only seems to work properly if `namedate: TRUE`. 
* **(author year) example:** This is a well known result [@ji20123d].

The bibliography will appear at the end of the document.

Though not normally available in the OUP LaTeX format, [CSL style files](https://www.zotero.org/styles) can also be used with the Rmarkdown adaptation by setting in the YAML header `citation_package: "default"` and defining the `csl` element to be the path towards the style file.

# Equations

An equation without a label for cross-referencing:

$$
E=mc^2
$$

An inline equation: $y=ax+b$

An equation with a label for cross-referencing:

\begin{equation}\label{eq:eq1}
\int^{r_2}_0 F(r,\varphi){\rm d}r\,{\rm d}\varphi = 1
\end{equation}

This equation can be referenced as follows: Eq. \ref{eq:eq1}

# Inserting R figures

The code below creates a figure.

```{r fig1, fig.cap="This is the first figure."}
plot(1:10,main="Some data",xlab="Distance (cm)",
     ylab="Time (hours)")
```

You can reference this figure as follows: Fig. \ref{fig:fig1}.

## Figures spanning two-columns

Figures can span two columns be setting `fig.env="figure*"`.

```{r fig2, fig.cap="This is a wide figure.", fig.env="figure*"}
plot(1:5,pch=19,main="Some wide data",
     xlab="Distance (cm)",ylab="Time (hours)")
```

Reference to second figure: Fig. \ref{fig:fig2}

# Tables

## Generate a table using `xtable`

```{r xtabletable, results="asis", echo=TRUE}
df <- data.frame(ID=1:3,code=letters[1:3])

# Creates tables that follow OUP guidelines 
# using xtable

print(xtable::xtable(df,caption="This is a xtable table.", label="tab:tab1"), comment=FALSE,caption.placement="top")
```

You can reference this table as follows: Table \ref{tab:tab1}.

## Generate a table using `kable`

```{r kabletable, echo=TRUE}
df <- data.frame(ID=1:3,code=letters[1:3])

# kable can alse be used for creating tables
knitr::kable(df,caption="This is a kable table.",
             booktabs=TRUE,label="tab2")
```

You can reference this table as follows: Table \ref{tab:tab2}.

## Table spanning two columns

Tables can span two columns be setting `table.envir = "table*"` in `knitr::kable`.

```{r widetable, echo=TRUE}
df <- data.frame(ID=1:3,code1=letters[1:3],
                code2=letters[4:6],
                code3=letters[7:9],
                code4=letters[10:12],
                code5=letters[13:15])

# kable can alse be used for creating tables
knitr::kable(df,caption="This is a wide kable table.",
             #format="latex",
             table.envir="table*",
             booktabs=TRUE,label="tab3")
```

# Cross-referencing sections

You can cross-reference sections and subsections as follows: Section \ref{literature-citations} and Section \ref{a-subsection}.

***Note:*** the last section in the document will be used as the section title for the bibliography.

For more portable and flexible referencing of sections, equations, figures and tables, use [`bookdown::pdf_document2`](https://github.com/rstudio/bookdown) with YAML header option `base_format: rticles::oup_article`.

# Appendices {-}

::: {.appendices latex=true}

# Section title of first appendix

blabla

## Subsection title of first appendix

and so on....
:::

# References

### 2022-01-09 manuscript.Rmd
Update manuscript.Rmd

@@
 
-bibliography: manuscript/mybibfile.bib # Includes refs in OUP example template
+bibliography: manuscript/bibliography.bib # Includes refs in OUP example template
 # When using `citation_package = "default"`, a CSL file can be used

### 2022-01-09 manuscript.Rmd
Update manuscript.Rmd

@@
 ---
-title: Template for Oxford University Press papers
+title: Investigating factors that may influence COVID-19 survival
 author:
-  - name: Alpha Beta
-    email: alpha@university.ac.uk
-    corresponding_author: TRUE # Automatic footnote using corresponding_author
+  - name: 36112895
+    email: 36112895@lancaster.ac.uk
+    corresponding_author: TRUE
     affiliation:
-      - ABC
-      - DEF
-    biography: |
-      A researcher.
-      #    biography_pic: "/PATH/TO/image.png"
-      #    biography_pic_width: "1in"
-  - name: Beta Gamma
-    email: beta@university.ac.uk
-    affiliation: DEF
-    footnote: 1 # Footnote using reference
-  - name: Zeta Epsilon
-    email: zeta@institution.ac.uk
-    affiliation: XYZ
+      - A
     footnote:
       - 1
-        - 2
 address:
-  - code: ABC
-    department: Department of Biology
-    organization: University
-  - code: DEF
-    address: University
-  - code: XYZ
-    address: Institution
+  - code: A
+    department: Lancaster Medical School
+    organization: University of Lancaster
 footnote:
   - code: 1
-    text: "Equal contribution"
-  - code: 2
-    text: "Current email address: [alpha@university.ac.uk](mailto:alpha@university.ac.uk)"
+    text: "Postgraduate Student"
 abstract:
@@
   - prognostics
-competing_interests: |
-  There are no competing interest.
-author_contributions: |
-  Equal contributions
-acknowledgements: |
-  This is an acknowledgement.
-
-  It consists of two paragraphs.
 

### 2022-01-09 R/modelling/ImputationStep.R
update R/modelling/ImputationStep.R

@@ -53,8 +53,8 @@ ImputationStep <- function (initial, phase, upload = TRUE) {
   processed <- ImputationProcessing(imputation = imputation, imputationdata = initial)
   
   
-  # if training
-  if (phase == 'training') {
+  # if training, or perhaps either training/testing
+  if (phase == 'training' | phase == 'testing') {
     # outcome date
     processed$outcome_date <- processed$admission_date + processed$time_to_outcome
 

### 2022-01-09 R/modelling/Testing.R
Update R/modelling/Testing.R

@@ -3,16 +3,16 @@
 # Created by: greyhypotheses
 # Created on: 06/01/2022
 
-Testing <- function (model, preliminary)  {
+Testing <- function (model, data)  {
 
   # Core: Narrow external validation
-  estimates <- survival::predict(object = model, newdata = preliminary)
+  estimates <- predict(object = model, newdata = data)
   # calibration plot
   # AUC
   # sensitivity analysis
 
   # Boosted
-  # mboost::survFit(object = model, newdata = preliminary )
+  # mboost::survFit(object = model, newdata = data)
   # calibration plot
   # AUC
   # sensitivity analysis

### 2022-01-09 R/modelling/interface.R
Update R/modelling/interface.R

@@ -53,11 +53,22 @@ training_ <- ImputationStep(initial = training[, variables],
 unboosted <- ModelCOXPH(training_ = training_, upload = FALSE)
 
 # testing
-testing_ <- ImputationStep(initial = testing[, variables[!(variables %in% dependent)]],
-                           phase = 'testing', upload = TRUE)
+evaluating_ <- testing[complete.cases(testing), ]
+testing_ <- ImputationStep(initial = testing[, variables],
+                           phase = 'testing', upload = FALSE)
 
-preliminary <- testing[complete.cases(testing), ]
-estimates <- predict(object = unboosted, newdata = preliminary, type = 'expected')
+
+
+
+# Play
+internal <- predict(object = unboosted, type = 'expected')
+internal <- data.frame(expected = internal)
+internal$lp <- predict(object = unboosted, type = 'lp')
+internal$risk_score <- predict(object = unboosted, type = 'risk')
+internal <- cbind(internal, as.data.frame(predict(object = unboosted, type = 'terms')))
+internal$survival_probability <- exp(-internal$expected)
+
+
 
 
 

### 2022-01-10 R/modelling/Testing.R
Delete R/modelling/Testing.R

@@ -1,20 +1,0 @@
-# Title     : Testing.R
-# Objective : Testing
-# Created by: greyhypotheses
-# Created on: 06/01/2022
-
-Testing <- function (model, data)  {
-
-  # Core: Narrow external validation
-  estimates <- predict(object = model, newdata = data)
-  # calibration plot
-  # AUC
-  # sensitivity analysis
-
-  # Boosted
-  # mboost::survFit(object = model, newdata = data)
-  # calibration plot
-  # AUC
-  # sensitivity analysis
-
-}

### 2022-01-10 R/evaluation/AssumptionViolationsCox.R
Add R/evaluation/AssumptionViolationsCox.R

@@ -1,0 +1,116 @@
+# Title     : AssumptionViolationsCox.R
+# Objective : Assumptions
+# Created by: greyhypotheses
+# Created on: 10/01/2022
+
+
+
+#' @param model: the training data model
+ViolationsProportionalityTable <- function (model) {
+
+  proportionality <- cox.zph(model)
+  estimates <- data.frame(proportionality$table)
+
+  return(estimates)
+
+}
+
+
+
+#' @param model: the training data model
+ViolationsTime <- function (model) {
+
+  proportionality <- cox.zph(model)
+
+  excerpt <- data.frame(proportionality$y) %>%
+    mutate(time = proportionality$time, x = proportionality$x)
+
+  excerpt %>%
+    select(age_group, malignant_neoplasm, time) %>%
+    gather(key = 'covariate', value = 'beta', -time) %>%
+    ggplot(mapping = aes(x = time, y = beta)) +
+    geom_point(alpha = 0.25) +
+    geom_smooth(method = 'loess', formula = 'y ~ x', size = 0.15, colour = 'black') +
+    xlim(NA, round(max(excerpt$time), digits = -2)) +
+    facet_wrap(~covariate, scales = 'free',
+               labeller = as_labeller(c('age_group' = 'Age Group', 'malignant_neoplasm' = 'Malignant Neoplasm'))) +
+    coord_trans(x = scales::pseudo_log_trans()) +
+    theme_minimal() +
+    theme(panel.spacing = unit(x = 3, units = 'lines'),
+          panel.grid.minor = element_blank(),
+          panel.grid.major = element_line(size = 0.1),
+          plot.title = element_text(hjust = 0.5),
+          strip.text.x = element_text(size = 11, face = 'bold'),
+          plot.caption = element_text(hjust = 0, size = 11, colour = 'darkgrey'),
+          axis.text.x = element_text(size = 11, angle = 90), axis.text.y = element_text(size = 11),
+          axis.title.x = element_text(size = 13), axis.title.y = element_text(size = 13)) +
+    xlab(label = '\ntime\n') +
+    ylab(label = '\ntime dependent coefficient\n')
+
+
+}
+
+
+
+#' @param data: training data
+ViolationsKaplan <- function (data) {
+
+  # the survfit strata builder
+  label <- function (text, times){
+    return(list(labels= data.frame(label = rep(text, times))))
+  }
+
+  # violation cases
+  cases <- list(age_group =
[…]

### 2022-01-10 R/evaluation/SurivalCurve.R
Add R/evaluation/SurivalCurve.R

@@ -1,0 +1,29 @@
+# Title     : SurivalCurve.R
+# Objective : SurvivalCurve
+# Created by: greyhypotheses
+# Created on: 10/01/2022
+
+
+#' @param data: training data
+SurvivalCurve <- function (data) {
+
+  null <- survfit(formula = Surv(time = data$time_to_outcome, event = data$deceased) ~ 1,
+                  data = data)
+  T <- data.frame(time = null$time, probability = null$surv, lower = null$lower, upper = null$upper)
+
+
+  ggplot(data = T, mapping = aes(x = time)) +
+    geom_ribbon(mapping = aes(ymin = lower, ymax = upper), fill = 'steelblue', alpha = 0.25) +
+    geom_line(mapping = aes(y = probability), alpha = 0.65, size = 0.5, linetype = 2, colour = 'black') +
+    theme_minimal() +
+    theme(panel.grid.minor = element_blank(),
+          panel.grid.major = element_line(size = 0.1),
+          plot.title = element_text(hjust = 0.5),
+          plot.caption = element_text(hjust = 0, vjust = 5, size = 11, colour = 'black'),
+          axis.text.x = element_text(size = 11, angle = 90), axis.text.y = element_text(size = 11),
+          axis.title.x = element_text(size = 13), axis.title.y = element_text(size = 13)) +
+    xlab(label = '\ntime\n') +
+    ylab(label = '\nprobability\n') +
+    labs(caption = str_wrap(string = 'Strata: All', width = 60) )
+
+}

### 2022-01-10 R/modelling/ModelCOXPH.R
Update R/modelling/ModelCOXPH.R

@@ -30,8 +30,7 @@ ModelCOXPH <- function (training_, upload = TRUE) {
     }
 
     # modelling
-    initial <- coxph(formula = formula , data = training_, model = TRUE)
-    unboosted <- step(object = initial)
+    unboosted <- coxph(formula = formula , data = training_, model = TRUE, x = TRUE, y = TRUE)
     save(unboosted, file = file.path(pathstr, 'unboosted'), ascii = TRUE, compress = TRUE, compression_level = 7)
 
   }

### 2022-01-10 R/modelling/interface.R
Update R/modelling/interface.R

@@ -52,21 +52,7 @@ training_ <- ImputationStep(initial = training[, variables],
                             phase = 'training', upload = TRUE)
 unboosted <- ModelCOXPH(training_ = training_, upload = FALSE)
 
-# testing
-evaluating_ <- testing[complete.cases(testing), ]
-testing_ <- ImputationStep(initial = testing[, variables],
-                           phase = 'testing', upload = FALSE)
-
-
-
-
-# Play
-internal <- predict(object = unboosted, type = 'expected')
-internal <- data.frame(expected = internal)
-internal$lp <- predict(object = unboosted, type = 'lp')
-internal$risk_score <- predict(object = unboosted, type = 'risk')
-internal <- cbind(internal, as.data.frame(predict(object = unboosted, type = 'terms')))
-internal$survival_probability <- exp(-internal$expected)
+# The Cox model inappropriate, consider univariate analysis for relevant covariates
 
 
 

### 2022-01-10 R/modelling/interface.R
Update R/modelling/interface.R

@@ -26,9 +26,6 @@ training <- dataframes$training
 testing <- dataframes$testing
 
 
-
-
-
 #' Imputation
 #'
 #'
@@ -42,17 +39,32 @@ variables <- c('admission_date', 'age_group', 'sex', 'asthma', 'liver_mild', 're
                'pulmonary', 'neurological', 'liver_mod_severe', 'malignant_neoplasm',
                'outcome', 'time_to_outcome')
 
-
 # exclude these variables during the testing phase
 dependent <- c('outcome', 'time_to_outcome')
 
-
-# training
+# for training
 training_ <- ImputationStep(initial = training[, variables],
                             phase = 'training', upload = TRUE)
+
+# for testing
+testing_ <- ImputationStep(initial = testing[, variables],
+                           phase = 'testing', upload = TRUE)
+
+
+#' Training
+#'
+#' The Cox model is inappropriate, however consider including univariate analysis for appropriate covariates
+#'
+#' For time depedent cases
+#' https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Time-dependent_covariate
+#' https://glmnet.stanford.edu/articles/Coxnet.html#cox-models-for-start-stop-data-1
+#' https://www.rdocumentation.org/packages/glmnet/versions/4.1-3/topics/glmnet
+#' https://www.rdocumentation.org/packages/survival/versions/3.2-13/topics/Surv
+#'
+#'
 unboosted <- ModelCOXPH(training_ = training_, upload = FALSE)
 
-# The Cox model inappropriate, consider univariate analysis for relevant covariates
+#
 
 
 

### 2022-01-15 R/missing/ImputationStep.R
Relocate R/modelling/ImputationStep.R -> R/missing/ImputationStep.R

@@ -6,7 +6,7 @@
 
 source(file = 'R/missing/Imputation.R')
 source(file = 'R/missing/ImputationProcessing.R')
-
+source(file = 'R/functions/StudyStart.R')
 
 
 #' ImputationStep
@@ -15,6 +15,8 @@ source(file = 'R/missing/ImputationProcessing.R')
 #' @param phase: training or testing
 #'
 ImputationStep <- function (initial, phase, upload = TRUE) {
+
+
 
 
   # ascertain phase value
@@ -67,6 +69,11 @@ ImputationStep <- function (initial, phase, upload = TRUE) {
     # the censor & deceased fields
     processed$censored <- dplyr::if_else(processed$outcome == 'Death', true = 0, false = 1)
     processed$deceased <- dplyr::if_else(processed$outcome == 'Death', true = 1, false = 0)
+
+    # study start point
+    tau <- StudyStart()
+    processed$start <- as.integer(processed$admission_date - tau)
+    processed$stop <- as.integer(processed$outcome_date - tau)
   }
 
 

### 2022-01-15 R/functions/StudyData.R
Update R/functions/StudyData.R

@@ -4,6 +4,7 @@
 # Created on: 30/12/2021
 
 source(file = 'R/functions/FrequenciesTable.R')
+source(file = 'R/functions/StudyStart.R')
 
 StudyData <- function () {
 
@@ -79,13 +80,17 @@ StudyData <- function () {
   data$time_to_outcome <- as.integer(data$outcome_date - data$admission_date)
 
   # implausible observations
-  condition <- (data$time_to_outcome < 0 & !is.na(data$time_to_outcome))
+  condition <- (data$time_to_outcome < 1 & !is.na(data$time_to_outcome))
   implausible <- data[condition,]
   plausible <- data[!condition,]
 
-
   # message
   print(paste('There are ', nrow(implausible), ' implausible observations; omitted.'))
+
+  # extra time variables
+  initial <- StudyStart()
+  plausible$start <- as.integer(plausible$admission_date - initial)
+  plausible$stop <- as.integer(plausible$outcome_date - initial)
 
 
   return(plausible)

### 2022-01-15 R/main.R
Update R/main.R

@@ -8,6 +8,7 @@
 #' Programs
 #'
 source(file = 'R/functions/StudyData.R')
+source(file = 'R/demographics/Is.R')
 
 
 
@@ -15,3 +16,11 @@ source(file = 'R/functions/StudyData.R')
 #'
 data <- StudyData()
 str(data)
+
+
+#' Is()
+#'
+#' Is the study's population demographics representative
+#' of England's population demographics
+#'
+Is()

### 2022-01-15 R/modelling/interface.R
Update R/modelling/interface.R

@@ -9,8 +9,11 @@
 #'
 source(file = 'R/functions/StudyData.R')
 source(file = 'R/functions/TemporalSplit.R')
-source(file = 'R/modelling/ImputationStep.R')
+source(file = 'R/missing/ImputationStep.R')
 source(file = 'R/modelling/ModelCOXPH.R')
+
+source(file = 'R/evaluation/AssumptionViolationsCox.R')
+source(file = 'R/evaluation/SurivalCurve.R')
 
 
 #' The data set
@@ -64,8 +67,10 @@ testing_ <- ImputationStep(initial = testing[, variables],
 #'
 unboosted <- ModelCOXPH(training_ = training_, upload = FALSE)
 
-#
-
+SurvivalCurve(data = training_)
+ViolationsKaplan(data = training_)
+ViolationsTime(model = unboosted)
+ViolationsProportionalityTable(model = unboosted)
 
 
 

### 2022-01-16 R/InstallPackages.R
Update R/InstallPackages.R

@@ -11,7 +11,7 @@ InstallPackages <- function (){
   # EpiEstim: Reproduction numbers of epidemics. (https://cran.r-project.org/web/packages/EpiEstim/index.html)
   packages <- c('tidyverse', 'data.table', 'ggplot2', 'rmarkdown', 'rstatix', 'latex2exp', 'moments', 'healthcareai',
                 'equatiomatic', 'survival', 'survminer', 'EpiEstim', 'rticles', 'ggcorrplot', 'rcompanion',
-                'patchwork', 'readxl', 'Hmisc', 'mice', 'matrixStats', 'devtools', 'caret', 'xgboost')
+                'patchwork', 'readxl', 'Hmisc', 'mice', 'matrixStats', 'devtools', 'caret', 'xgboost', 'survRM2')
 
   # Install
   .install <- function(x){

### 2022-01-16 R/functions/StudyData.R
Update R/functions/StudyData.R

@@ -69,6 +69,10 @@ StudyData <- function () {
   data$censored <- factor(x = data$censored, levels = c(1, 0), exclude = NULL)
   table(data$outcome, data$censored)
 
+  data$deceased <- dplyr::if_else(data$outcome == 'Death', true = 1, false = 0)
+  data$deceased <- factor(x = data$deceased, levels = c(0, 1), exclude = NULL)
+  table(data$outcome, data$deceased)
+
   # outcome date
   data$outcome_date <- as.Date(data$outcome_date)
   if (anyNA(data$outcome_date)) {

### 2022-01-16 R/evaluation/SurivalCurve.R
Delete R/evaluation/SurivalCurve.R

@@ -1,29 +1,0 @@
-# Title     : SurivalCurve.R
-# Objective : SurvivalCurve
-# Created by: greyhypotheses
-# Created on: 10/01/2022
-
-
-#' @param data: training data
-SurvivalCurve <- function (data) {
-
-  null <- survfit(formula = Surv(time = data$time_to_outcome, event = data$deceased) ~ 1,
-                  data = data)
-  T <- data.frame(time = null$time, probability = null$surv, lower = null$lower, upper = null$upper)
-
-
-  ggplot(data = T, mapping = aes(x = time)) +
-    geom_ribbon(mapping = aes(ymin = lower, ymax = upper), fill = 'steelblue', alpha = 0.25) +
-    geom_line(mapping = aes(y = probability), alpha = 0.65, size = 0.5, linetype = 2, colour = 'black') +
-    theme_minimal() +
-    theme(panel.grid.minor = element_blank(),
-          panel.grid.major = element_line(size = 0.1),
-          plot.title = element_text(hjust = 0.5),
-          plot.caption = element_text(hjust = 0, vjust = 5, size = 11, colour = 'black'),
-          axis.text.x = element_text(size = 11, angle = 90), axis.text.y = element_text(size = 11),
-          axis.title.x = element_text(size = 13), axis.title.y = element_text(size = 13)) +
-    xlab(label = '\ntime\n') +
-    ylab(label = '\nprobability\n') +
-    labs(caption = str_wrap(string = 'Strata: All', width = 60) )
-
-}

### 2022-01-16 R/functions/StudyData.R
Update R/functions/StudyData.R

@@ -66,11 +66,9 @@ StudyData <- function () {
   # right censored: 1, 0
   # outcome: alive, deceased
   data$censored <- dplyr::if_else(data$outcome == 'Death', true = 0, false = 1)
-  data$censored <- factor(x = data$censored, levels = c(1, 0), exclude = NULL)
   table(data$outcome, data$censored)
 
   data$deceased <- dplyr::if_else(data$outcome == 'Death', true = 1, false = 0)
-  data$deceased <- factor(x = data$deceased, levels = c(0, 1), exclude = NULL)
   table(data$outcome, data$deceased)
 
   # outcome date

### 2022-01-16 R/modelling/interface.R
Update R/modelling/interface.R

@@ -4,61 +4,30 @@
 # Created on: 06/01/2022
 
 
-
 #' Programs
 #'
-source(file = 'R/functions/StudyData.R')
-source(file = 'R/functions/TemporalSplit.R')
-source(file = 'R/missing/ImputationStep.R')
+source(file = 'R/functions/ImputedData.R')
 source(file = 'R/modelling/ModelCOXPH.R')
-
 source(file = 'R/evaluation/AssumptionViolationsCox.R')
-source(file = 'R/evaluation/SurivalCurve.R')
+source(file = 'R/evaluation/SurvivalCurve.R')
 
 
-#' The data set
+#' Data
 #'
-data <- StudyData()
-str(data)
-
-
-#' Splitting
+#' Upload the study data, excluding implausible observations, and
+#' its imputed form
 #'
-dataframes <- TemporalSplit(data = data)
-training <- dataframes$training
-testing <- dataframes$testing
-
-
-#' Imputation
-#'
-#'
-#' ...  re-calculate: censored & outcome_date, and add deceased
-#' ...  the imputation variables exclude: outcome_date (date can't be used
-#'      in the MICE models, time_to _outcome in lieu), censored (it is a project
-#'      variable created for analysis & graphing purposes)
-
-# all variables for training phase
-variables <- c('admission_date', 'age_group', 'sex', 'asthma', 'liver_mild', 'renal',
-               'pulmonary', 'neurological', 'liver_mod_severe', 'malignant_neoplasm',
-               'outcome', 'time_to_outcome')
-
-# exclude these variables during the testing phase
-dependent <- c('outcome', 'time_to_outcome')
-
-# for training
-training_ <- ImputationStep(initial = training[, variables],
-                            phase = 'training', upload = TRUE)
-
-# for testing
-testing_ <- ImputationStep(initial = testing[, variables],
-                           phase = 'testing', upload = TRUE)
+dataframes <- ImputedData(upload = TRUE)
+data <- dataframes$data
+training_ <- dataframes$training_
+testing_ <- dataframes$testing_
 
 
 #' Training
 #'
-#' The Cox model is inappropriate, however consider including univariate analysis for appropriate covariates
+#' The Cox model is inappropriate, proportionality assumption does not hold.
 #'
-#' For time depedent cases
+#' For time depedent cases:
[…]

### 2022-01-16 R/evaluation/KaplanMeier.R
Add R/evaluation/KaplanMeier.R

@@ -1,0 +1,40 @@
+# Title     : KaplanMeier.R
+# Objective : Kaplan Meier
+# Created by: greyhypotheses
+# Created on: 16/01/2022
+
+
+
+#' Programs
+#'
+source(file = 'R/functions/ImputedData.R')
+source(file = 'R/evaluation/SurvivalCurve.R')
+
+
+KaplanMeier <- function () {
+
+  # Upload the study data, excluding implausible observations, and its imputed form
+  dataframes <- ImputedData(upload = TRUE)
+  data <- dataframes$data
+  training_ <- dataframes$training_
+  testing_ <- dataframes$testing_
+
+  # complete case
+  complete <- data %>%
+    dplyr::filter(!is.na(time_to_outcome) & !is.na(deceased))
+
+  # imputed
+  imputed <- dplyr::bind_rows(training_, testing_)
+
+  # the survival curves
+  left <- SurvivalCurve(
+    data = complete,
+    caption = str_glue('A. {nrow(complete)} records'))
+  right <- SurvivalCurve(
+    data = imputed,
+    caption = str_glue('B. {nrow(imputed)} records'))
+
+  left + right
+
+}
+

### 2022-01-16 R/functions/ImputedData.R
Add R/functions/ImputedData.R

@@ -1,0 +1,44 @@
+# Title     : ImputedData.R
+# Objective : Imputed Data
+# Created by: greyhypotheses
+# Created on: 16/01/2022
+
+source(file = 'R/functions/StudyData.R')
+source(file = 'R/functions/TemporalSplit.R')
+source(file = 'R/missing/ImputationStep.R')
+
+ImputedData <- function(upload = TRUE) {
+
+  # The data set
+  data <- StudyData()
+
+
+  # Splitting
+  dataframes <- TemporalSplit(data = data)
+  training <- dataframes$training
+  testing <- dataframes$testing
+  rm(dataframes)
+
+
+  # all variables for training phase
+  variables <- c('admission_date', 'age_group', 'sex', 'asthma', 'liver_mild', 'renal',
+                 'pulmonary', 'neurological', 'liver_mod_severe', 'malignant_neoplasm',
+                 'outcome', 'time_to_outcome')
+
+
+  # imputation
+  #
+  #
+  # ...  re-calculate: censored & outcome_date, and add deceased
+  # ...  the imputation variables exclude: outcome_date (date can't be used
+  #      in the MICE models, time_to _outcome in lieu), censored (it is a project
+  #      variable created for analysis & graphing purposes)
+  training_ <- ImputationStep(initial = training[, variables],
+                              phase = 'training', upload = upload)
+
+  testing_ <- ImputationStep(initial = testing[, variables],
+                             phase = 'testing', upload = upload)
+
+  return(list(data = data, training_ = training_, testing_ = testing_))
+
+}

### 2022-01-16 R/evaluation/KaplanMeier.R
Update R/evaluation/KaplanMeier.R

@@ -7,7 +7,7 @@
 
 #' Programs
 #'
-source(file = 'R/functions/ImputedData.R')
+source(file = 'R/functions/ExtensiveStudyData.R')
 source(file = 'R/evaluation/SurvivalCurve.R')
 
 

### 2022-01-16 R/functions/ExtensiveStudyData.R
Update R/functions/ExtensiveStudyData.R

@@ -1,4 +1,4 @@
-# Title     : ImputedData.R
+# Title     : ExtensiveStudyData.R
 # Objective : Imputed Data
 # Created by: greyhypotheses
 # Created on: 16/01/2022

### 2022-01-16 R/functions/StudyData.R
Update R/functions/StudyData.R

@@ -72,22 +72,17 @@ StudyData <- function () {
   table(data$outcome, data$deceased)
 
   # outcome date
+  # anyNA(data$outcome_date)
   data$outcome_date <- as.Date(data$outcome_date)
-  if (anyNA(data$outcome_date)) {
-    missing <- sum(is.na(data$outcome_date))
-    print(paste0('There are missing outcome dates: ', missing, ' missing'))
-  }
+
 
   # time_to_event
   data$time_to_outcome <- as.integer(data$outcome_date - data$admission_date)
 
   # implausible observations
   condition <- (data$time_to_outcome < 1 & !is.na(data$time_to_outcome))
-  implausible <- data[condition,]
   plausible <- data[!condition,]
 
-  # message
-  print(paste('There are ', nrow(implausible), ' implausible observations; omitted.'))
 
   # extra time variables
   initial <- StudyStart()

### 2022-01-16 R/modelling/interface.R
Update R/modelling/interface.R

@@ -6,7 +6,7 @@
 
 #' Programs
 #'
-source(file = 'R/functions/ImputedData.R')
+source(file = 'R/functions/ExtensiveStudyData.R')
 source(file = 'R/modelling/ModelCOXPH.R')
 source(file = 'R/evaluation/AssumptionViolationsCox.R')
 source(file = 'R/evaluation/SurvivalCurve.R')

### 2022-01-17 R/evaluation/KaplanMeier.R
Update R/evaluation/KaplanMeier.R

@@ -14,7 +14,7 @@ source(file = 'R/evaluation/SurvivalCurve.R')
 KaplanMeier <- function () {
 
   # Upload the study data, excluding implausible observations, and its imputed form
-  dataframes <- ImputedData(upload = TRUE)
+  dataframes <- ExtensiveStudyData(upload = TRUE)
   data <- dataframes$data
   training_ <- dataframes$training_
   testing_ <- dataframes$testing_

### 2022-01-17 R/functions/ExtensiveStudyData.R
Update R/functions/ExtensiveStudyData.R

@@ -7,7 +7,7 @@ source(file = 'R/functions/StudyData.R')
 source(file = 'R/functions/TemporalSplit.R')
 source(file = 'R/missing/ImputationStep.R')
 
-ImputedData <- function(upload = TRUE) {
+ExtensiveStudyData <- function(upload = TRUE) {
 
   # The data set
   data <- StudyData()
@@ -39,6 +39,9 @@ ImputedData <- function(upload = TRUE) {
   testing_ <- ImputationStep(initial = testing[, variables],
                              phase = 'testing', upload = upload)
 
-  return(list(data = data, training_ = training_, testing_ = testing_))
+  data_ <- ImputationStep(initial = data[, variables],
+                          phase = 'data', upload = upload)
+
+  return(list(data = data, training_ = training_, testing_ = testing_, data_ = data_))
 
 }

### 2022-01-17 R/main.R
Update R/main.R

@@ -8,6 +8,7 @@
 #' Programs
 #'
 source(file = 'R/functions/StudyData.R')
+source(file = 'R/functions/ExtensiveStudyData.R')
 source(file = 'R/demographics/Is.R')
 
 
@@ -16,6 +17,11 @@ source(file = 'R/demographics/Is.R')
 #'
 data <- StudyData()
 str(data)
+
+dataframes <- ExtensiveStudyData(upload = TRUE)
+training_ <- dataframes$training_
+testing_ <- dataframes$testing_
+data_ <- dataframes$data_
 
 
 #' Is()

### 2022-01-17 R/missing/ImputationStep.R
Update R/missing/ImputationStep.R

@@ -20,7 +20,7 @@ ImputationStep <- function (initial, phase, upload = TRUE) {
 
 
   # ascertain phase value
-  if (!(phase %in% c('training', 'testing'))) {
+  if (!(phase %in% c('training', 'testing', 'data'))) {
     stop("The phase parameter must be a string equal to 'training' or 'testing'")
   }
 
@@ -56,7 +56,7 @@ ImputationStep <- function (initial, phase, upload = TRUE) {
   
   
   # if training, or perhaps either training/testing
-  if (phase == 'training' | phase == 'testing') {
+  if (phase %in% c('training', 'testing', 'data')) {
     # outcome date
     processed$outcome_date <- processed$admission_date + processed$time_to_outcome
 

### 2022-01-17 R/modelling/interface.R
Update R/modelling/interface.R

@@ -17,7 +17,7 @@ source(file = 'R/evaluation/SurvivalCurve.R')
 #' Upload the study data, excluding implausible observations, and
 #' its imputed form
 #'
-dataframes <- ImputedData(upload = TRUE)
+dataframes <- ExtensiveStudyData(upload = TRUE)
 data <- dataframes$data
 training_ <- dataframes$training_
 testing_ <- dataframes$testing_

### 2022-01-18 R/evaluation/KaplanMeier.R
Update R/evaluation/KaplanMeier.R

@@ -16,23 +16,19 @@ KaplanMeier <- function () {
   # Upload the study data, excluding implausible observations, and its imputed form
   dataframes <- ExtensiveStudyData(upload = TRUE)
   data <- dataframes$data
-  training_ <- dataframes$training_
-  testing_ <- dataframes$testing_
+  data_ <- dataframes$data_
 
   # complete case
   complete <- data %>%
     dplyr::filter(!is.na(time_to_outcome) & !is.na(deceased))
-
-  # imputed
-  imputed <- dplyr::bind_rows(training_, testing_)
 
   # the survival curves
   left <- SurvivalCurve(
     data = complete,
     caption = str_glue('A. {nrow(complete)} records'))
   right <- SurvivalCurve(
-    data = imputed,
-    caption = str_glue('B. {nrow(imputed)} records'))
+    data = data_,
+    caption = str_glue('B. {nrow(data_)} records'))
 
   left + right
 

### 2022-01-18 R/main.R
Update R/main.R

@@ -7,21 +7,22 @@
 
 #' Programs
 #'
-source(file = 'R/functions/StudyData.R')
 source(file = 'R/functions/ExtensiveStudyData.R')
 source(file = 'R/demographics/Is.R')
 
 
 
-#' The data set
+#' The data
 #'
-data <- StudyData()
-str(data)
-
 dataframes <- ExtensiveStudyData(upload = TRUE)
+data <- dataframes$data
 training_ <- dataframes$training_
 testing_ <- dataframes$testing_
 data_ <- dataframes$data_
+rm(dataframes)
+
+str(data)
+
 
 
 #' Is()

### 2022-01-18 R/modelling/ModelCOXPH.R
Delete R/modelling/ModelCOXPH.R

@@ -1,41 +1,0 @@
-# Title     : ModelCOXPH.R
-# Objective : survival::coxph
-# Created by: greyhypotheses
-# Created on: 09/01/2022
-
-ModelCOXPH <- function (training_, upload = TRUE) {
-
-  # the model object directory path
-  pathstr <- file.path(getwd(), 'warehouse', 'training', 'models')
-  if (!dir.exists(paths = pathstr)) {
-    dir.create(path = pathstr, showWarnings = TRUE, recursive = TRUE)
-  }
-
-
-  # formula
-  formula <- Surv(time = training_$time_to_outcome, event = training_$deceased) ~ age_group + sex + asthma +
-    liver_mild + renal + pulmonary + neurological + liver_mod_severe + malignant_neoplasm
-
-
-  # upload or run
-  if (upload) {
-
-    load(file.path(pathstr, 'unboosted'))
-
-  } else {
-
-    # a directory for the resulting models
-    if (file.exists(paths = file.path(pathstr, 'unboosted'))) {
-      base::unlink(file.path(pathstr, 'unboosted'), recursive = TRUE)
-    }
-
-    # modelling
-    unboosted <- coxph(formula = formula , data = training_, model = TRUE, x = TRUE, y = TRUE)
-    save(unboosted, file = file.path(pathstr, 'unboosted'), ascii = TRUE, compress = TRUE, compression_level = 7)
-
-  }
-
-  return(unboosted)
-
-
-}

### 2022-01-18 R/modelling/interface.R
Update R/modelling/interface.R

@@ -7,7 +7,6 @@
 #' Programs
 #'
 source(file = 'R/functions/ExtensiveStudyData.R')
-source(file = 'R/modelling/ModelCOXPH.R')
 source(file = 'R/evaluation/AssumptionViolationsCox.R')
 source(file = 'R/evaluation/SurvivalCurve.R')
 
@@ -21,6 +20,20 @@ dataframes <- ExtensiveStudyData(upload = TRUE)
 data <- dataframes$data
 training_ <- dataframes$training_
 testing_ <- dataframes$testing_
+data_ <- dataframes$data_
+
+even <- data[complete.cases(data),]
+
+source(file = 'R/modelling/Univariate.R')
+analysis <- Univariate(blob = training_)
+analysis <- Univariate(blob = data_)
+analysis <- Univariate(blob = even)
+
+
+source(file = 'R/modelling/Multivariate.R')
+analysis <- Multivariate(blob = training_)
+analysis <- Multivariate(blob = data_)
+analysis <- Multivariate(blob = even)
 
 
 #' Training
@@ -34,7 +47,7 @@ testing_ <- dataframes$testing_
 #' https://www.rdocumentation.org/packages/survival/versions/3.2-13/topics/Surv
 #'
 #'
-unboosted <- ModelCOXPH(training_ = training_, upload = FALSE)
+
 
 SurvivalCurve(data = training_, caption = 'Training Data (multiply imputed)')
 ViolationsKaplan(data = training_)

### 2022-01-18 R/modelling/Multivariate.R
Add R/modelling/Multivariate.R

@@ -1,0 +1,44 @@
+# Title     : Multivariate.R
+# Objective : Multivariate
+# Created by: greyhypotheses
+# Created on: 17/01/2022
+
+
+Multivariate <- function(blob) {
+  
+  # formula
+  formula <- Surv(time = blob$time_to_outcome, event = blob$deceased) ~ age_group + sex + asthma +
+    liver_mild + renal + pulmonary + neurological + liver_mod_severe + malignant_neoplasm
+
+
+  # model
+  model <- coxph(formula = formula , data = blob, model = TRUE, x = TRUE, y = TRUE)
+
+  # diagnostics of coefficients
+  obj <- summary(object = model)
+  C <- data.frame(obj$coefficients)
+  interval <- data.frame(obj$conf.int)
+
+  # proportionality
+  proportionality <- cox.zph(fit = model)
+  evidence <- data.frame(proportionality$table[, 'p'])
+  names(evidence) <- 'proportionality'
+  row.names(evidence) <- c('age_group', 'sexFemale', 'asthmayes', 'liver_mildyes', 'renalyes',
+                           'pulmonaryyes', 'neurologicalyes', 'liver_mod_severeyes', 'malignant_neoplasmyes', 'GLOBAL')
+
+  # summarise
+  frame <- cbind(C[, c('coef', 'exp.coef.')], interval[, c('lower..95', 'upper..95')]) %>%
+    mutate(p_value = C[, 'Pr...z..'])
+  frame <- base::merge(x = frame, y = evidence, by = 0, all.x = TRUE)
+
+  # age group
+  indices <- startsWith(frame$Row.names, prefix = 'age_group')
+  frame[indices, 'proportionality'] <- evidence[row.names(evidence) == 'age_group',]
+
+  # field names
+  names(frame) <- c('Factor', 'Coefficient', 'HR', 'HRLCI', 'HRUCI', 'p_value', 'proportionality')
+
+  return(list(diagnostics = frame, model = model))
+
+}
+

### 2022-01-18 R/modelling/Univariate.R
Add R/modelling/Univariate.R

@@ -1,0 +1,56 @@
+# Title     : Univariate.R
+# Objective : Univariate
+# Created by: greyhypotheses
+# Created on: 17/01/2022
+
+
+Univariate <- function (blob) {
+
+  # the fields
+  fields <- c( 'age_group', 'sex', 'asthma', 'liver_mild', 'renal', 'pulmonary',
+               'neurological', 'liver_mod_severe', 'malignant_neoplasm')
+
+  # a formula per prospective predictor
+  formulae <- sapply(fields,
+                     function (x) as.formula(paste('Surv(time = time_to_outcome, event = deceased) ~', x)))
+
+  # the cox model of each
+  models <- lapply(X = formulae, FUN = function(formula){coxph(formula = formula, data = blob)})
+
+  # a function for extracting diagnostics of interest
+  estimates <- function (model) {
+
+    # a model
+    obj <- summary(object = model)
+    
+    # the coefficient diagnostics
+    C <- obj$coefficients
+    indices <- rownames(C)
+    interval <- obj$conf.int
+
+    # proportionality assumption
+    proportionality <- cox.zph(fit = model)
+    variate <- names(model$xlevels)
+    evidence <- proportionality$table[variate, 'p']
+
+    # summarising
+    frame <- data.frame(Coefficient = C[, 'coef'], HR = C[, 'exp(coef)'],
+                        HRLCI = interval[, 'lower .95'], HRUCI = interval[, 'upper .95'],
+                        p_value = C[, 'Pr(>|z|)'], proportionality = evidence)
+    row.names(frame) <- indices
+
+    return(frame)
+    
+  }  
+  est <- lapply(X = models, FUN = function (model){estimates(model = model)})
+  diagnostics <- dplyr::bind_rows(est)
+  
+  return(list(diagnostics = diagnostics, models = models))
+}
+
+
+
+
+
+
+

### 2022-01-18 paper.Rmd
Add paper.Rmd

---
title: Investigating factors that may influence COVID-19 survival
author:
  - name: 36112985
    email: 36112985@lancaster.ac.uk
    affiliation: Lancaster Medical School
    corresponding: 36112985@lancaster.ac.uk
    affiliation: 
      - Lancaster Medical School
address:
  - code: Lancaster Medical School
    address: Lancaster Medical School, Faculty of Health & Medicine, University of Lancaster
abstract: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

author_summary: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

bibliography: paper/mybibfile.bib
output:
  rticles::plos_article:
    keep_tex: FALSE
# plos.csl can be download and used locally
csl: http://www.zotero.org/styles/plos
---

_Text based on plos sample manuscript, see <https://journals.plos.org/ploscompbiol/s/latex>_

# Introduction

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

A list

- Item 1
- Item 2

Here are two sample references: @Feynman1963118 [@Dirac1953888].

# Methods

## Exploratory Data Analysis

* Relationships
* Correlations

## Missing Data

Most of the data's variables have missing values.  Prior to deciding how to address missing values, it is important to understand the missing values patterns [@steyerberg2010]. Rubin [@rubin1976, @steyerberg2010, @little2019] outlines three fundamental missing data mechanisms

* missing completely at random (MCAR) / administrative errors, accidents
* missing at random (MAR) / missing data associated with known patient characteristics [independent variables], or the outcome
* missing not at random (MNAR) / missing data associated with missing values of the factor/predictor in question or with unobserved predictors

If the missing values of a data set in question are *missing completely at random* then complete case analysis will suffice because the complete case excerpt is akin to a random sample from a complete population.  If MCAR does not hold, e.g., data is *missing at random*, then the complete case excerpt is not representative of the underlying population, therefore population inference is not possible via complete case analysis. [@steyerberg2010]

Herein, missing data mechanisms analysis is via logistic regression.  Almost all the independent variables have missing values patterns that violate MCAR; addressed via multiple imputation in-line with the advice and protocols of [@sterne2009] & [@steyerberg2010].

## Null Survival Curve

* Complete Case
* Imputation Case

## Uni-variate Analysis

* The table ... Cox

## Adjusted Analysis via Restricted Mean Survival Time

The proportional hazard assumption does not hold ...

# Results

# Discussion

Within a missing values setting, there are 2 options w.r.t. predictor effect analysis [@steyerberg2010]

* complete case analysis for the uni-variate analysis, and complete case predictors & imputed confounding variables for adjusted analysis.
* uni-variate and adjusted analysis after imputation of all missing values.

The investigation opted for the latter, especially because ...

## Bias

Possible bias points

* By virtue of the study's design pre-study--admission differences between hospitals is possible due to the ambiguity of the study's patient recruitment terms, and this might lead to post-study differences.  In a nutshell, the study is susceptible to selection bias. [Trochim]

* Missing data [@sterne2009]

# Conclusion

The analysis of missing values hints at flaws in data collection.

# References {#references .unnumbered}

### 2022-01-18 manuscript.Rmd
Update manuscript.Rmd

@@
 title: Investigating factors that may influence COVID-19 survival
-author:
-  - name: 36112895
-    email: 36112895@lancaster.ac.uk
-    corresponding_author: TRUE
-    affiliation:
-      - A
-    footnote:
-      - 1
-address:
-  - code: A
-    department: Lancaster Medical School
-    organization: University of Lancaster
-footnote:
-  - code: 1
-    text: "Postgraduate Student"
+author: 36112985
 abstract:
+  - "**Background:**"
   - "**Objective:**  To Investigate, and identify, the factors that may influence patient survival of coronavirus disease 19."
-  - "More"
+  - "**Methods**"
   - "**Motivation:** You can also have some paragraphs start with bold face."
-keywords:
-  - mathematics
-  - survival
-  - prognostics
-
-bibliography: manuscript/bibliography.bib # Includes refs in OUP example template
-# When using `citation_package = "default"`, a CSL file can be used
-# csl: https://www.zotero.org/styles/oxford-university-press-note
+bibliography: manuscript/bibliography.bib
 output:
-  rticles::oup_article:
-    oup_version: 1 # 1 = 2020 CTAN OUP CLS package; 0 = 2009 OUP CLS package
+  rmarkdown::pdf_document:
     keep_tex: FALSE
-    #journal: "Journal Name"
-    document_style: "contemporary" # Can be contemporary, modern, traditional
-    papersize: "large" # Can be large, medium, small
-    #citation_package: "default" # Uncomment when using a CSL; default "natbib"
-    namedate: TRUE # Set FALSE to use numeric refs; Default FALSE
-    #number_sections: FALSE # Uncomment to not number sections; default TRUE
-    #number_lines: TRUE # Use `lineno` package - Default FALSE
-    #number_lines_options: ["mathlines","switch"]  # Options for latex lineno package.
-    #onecolumn: TRUE # Uncomment for one column format; default FALSE
+    citation_package: "default"
+    number_sections: FALSE
     extra_dependencies:
-      - booktabs # to use with knitr::kable() example below
-
-# Example of pandoc's variable
-# urlcolor: orange
-# linkcolor: green
-# citecolor: red
+      - booktabs
+urlcolor: orange
+linkcolor: green
+citecolor: red
 header-includes:
-  - \theoremstyle{thmstyleone}
+  - \usepackage{color}
+  - \usepackage{caption}
+  - \captionsetup[figure]{font={normalsize, color=gray}, width=.8\linewidth}
+  - \usepackage{xcolor}
+  - \usepackage{titlesec}
+  - \usepackage{fancyhdr}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{proposition}[theorem]{Proposition}
-  - \theoremstyle{thmstyletwo}
-  - \newtheorem{example}{Example}
   - \newtheorem{remark}{Remark}
-  - \theoremstyle{thmstylethree}
   - \newtheorem{definition}{Definition}
+  - \definecolor{darkgrey}{HTML}{4a4a4a}
+  - \titleformat*{\section}{\Large\color{darkgrey}}
+  - \pagestyle{fancy}
 ---
@@
 --->
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, fig.pos = 'th',
-  out.width = '100%', dpi = 300, fig.env="figure")
-
-options(knitr.table.format = "latex") # For kable tables to write LaTeX table directly
+```{r include=FALSE}
+knitr::opts_chunk$set(echo = FALSE)
+options(knitr.table.format = "latex")
 ```
@@
 library(xtable)
+library(ggplot2)
 library(dplyr)
+library(tidyr)
+library(stringr)
+
 library(magrittr)
+library(latex2exp)
+library(survival)
+library(patchwork)
 ```
@@
 ```{r}
+sys.source(file = 'R/evaluation/KaplanMeier.R', envir = knitr::knit_global())
 sys.source(file = 'R/functions/StudyData.R', envir = knitr::knit_global())
@@
 
-This template is based on the generic OUP authoring template available on CTAN under [oup-authoring-template](https://www.ctan.org/pkg/oup-authoring-template).
-The CTAN template includes LaTeX documentation and a sample LaTeX document that provide far more details regarding the full functionality of the format.
-Here, only the basic functioning of the Rmarkdown adaptation of the format is demonstrated.
+During the past 2years, infections due to severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2) have led to a high number of coronavirus 19 disease deaths and in-hospital care.
+With time, case fatality rates, and in-hospital survival rates, have improved [...] as our understanding of coronavirus 19 disease pathogenesis has developed, and hence prevention, mitigation, and treatment measures have improved.
+Effective disease treatment measures depend on our understanding of the disease's prognosis factors [@riley2013], the focus of prognostic factors research.
+Prognostic factor research *aims to identify factors associated with subsequent clinical outcome in people with a particular disease or health condition* [@riley2013].
 
-## A subsection
-
-A numbered list:
-
-1) First point
-2) Second point
-    - Subpoint
-
-A bullet list:
-
-* First point
-* Second point
-
-## Notes
-
-* Extra white space in document will tend to disappear as text is filled in.
-* Code blocks tend to generate lots of empty white space when `echo=TRUE` for some reason.
+This brief investigates the prognostic factor potential of a set of measures associated with a prospective cohort study.
+The study's hypothesis is that *one or more of the measures has a significant impact on each patient's survival probability*.
 
@@
 
-## Data Source
-The data is a derivative of the ... It is a cohort study wherein ...
+## Study design and setting
+
+This investigation's data is an indirect ISARIC (International Severe Acute Respiratory Infection Consortium)  4C (Coronavirus Clinical Characterisation Consortium) study data set.
+The study is an ongoing prospective cohort study based in more than 200 United Kingdom acute care hospitals.
+This investigation's data is an England only data set.
+The reporting ...
+
+\vspace{20pt}
 
@@
 
-* Study setting
-* Eligibility criteria
+Inclusion criteria were people of all ages who were admitted to one of 208 acute care hospitals in England, Scotland, and Wales with proven or high likelihood of infection with a pathogen of public health interest, defined as SARS-CoV-2 for this event by Public Health England.
+Reverse transcriptase polymerase chain reaction was the only mode of testing available during the period of study.
+The decision to test was at the discretion of the clinician attending the patient, and not defined by protocol.
+The enrolment criterion “high likelihood of infection” reflects that a preparedness protocol cannot assume that a diagnostic test will be available for an emergent pathogen.
+Site training emphasises that only patients who tested positive for covid-19 were eligible for enrolment.
 
-## Outcome
+National guidance was provided by Public Health England and other UK public health agencies that advised who to test based on clinical case definitions for possible covid-19.
+We also included patients who had been admitted for a separate condition but had tested positive for covid-19 during their hospital stay.
+Patients were only enrolled during their index admission.
+Patients had clinical information from their routine health records uploaded into the case report form.
+Consent was not required for collection of depersonalised routine healthcare data for research in England and Wales.
+A waiver for consent was given by the Public Benefit and Privacy Panel in Scotland.
 
-* Died
+\vspace{20pt}
 
-## Predictors
+## Data
 
-# Literature citations
+Although the consortium collects a large variety of demographic, clinical, and outcome data, this investigation's data only included a limited subset of the study's variables.
+The demographic data variables are sex and age group, the comorbidity variables are asthma (physician diagnosed), mild liver disease, moderate/severe liver disease, renal disease, chronic pulmonary disease (excluding asthma), chronic neurological disorder, malignant neoplasm.
+
+The ISARIC 4C study collects its baseline data via paper case report forms developed by ISARIC, and the WHO (World Health Organization), for outbreak investigations.
+Additionally, the data brief states that the baseline data records are *uploaded [on] admission, and usually before hospital episodes were complete, to a REDCap database (Research Electronic Data Capture, Vanderbilt University, US, hosted by University of Oxford, UK).*
+
+\vspace{20pt}
+
+## Outcomes
+
+The ISARIC 4C study outcomes are: discharged alive, remains in hospital, death, transferred, and palliative discharge. Herein, the overarching outcomes of interest are: dead, alive.
+
+\vspace{20pt}
+
+## Independent Variables
+
+The predictors, factors, under investigation
+
+\vspace{20pt}
+
+## Factor Analysis
+
+The Cox Hazard, Restricted Mean Survival
+
+\vspace{20pt}
+
+# Results
+
+## Characteristics
+
+\begin{table}[h!]
+    \centering
+    \renewcommand{\arraystretch}{1.25}
+    \begin{tabular}{l r r r}
+    & & & \\
+    \textbf{In Hospital Mortality} & & & \\
+    \hline
+     & \textbf{no} & \textbf{yes} & \textbf{NA} \\
+    deceased & 31724 & 14049 & 347 \\
+     & 0.688 & 0.305 & 0.008 \\
+     & & & \\
+    \textbf{Comorbidities} & & & \\
+    \hline
+     & \textbf{no} & \textbf{yes} & \textbf{NA} \\
+    asthma & 37376 & 5872 & 2872 \\
+    & 0.81 & 0.13 & 0.06 \\
+    mild liver disease & 42021 & 666 & 3433 \\
+     & 0.91 & 0.01 & 0.07 \\
+    renal disease & 35158 & 8056 & 2906 \\
+     & 0.76 & 0.17 & 0.06 \\
+    pulmonary & 35265 & 8078 & 2777 \\
+     & 0.76 & 0.18 & 0.06 \\
+    neurological disorder & 37227 & 5806 & 3087 \\
+     & 0.81 & 0.13 & 0.07 \\
+    liver mod severe &  41901 &  902 & 3317 \\
+     & 0.91 & 0.02 & 0.07 \\
+    malignant neoplasm & 38164 & 4734 & 3222 \\
+     & 0.83 & 0.10 & 0.07 \\
+     & & & \\
+    \textbf{Demographics} & & & \\
+    \hline
+    \textbf{Age Group} &  \textbf{Male} & \textbf{Female} & \textbf{NA} \\
+    80-89 & 7067 & 6197 & 31 \\
+    & 0.153 & 0.134 & 0.001 \\
+    70-79 & 6388 & 4445 & 19 \\
+    & 0.139 & 0.096 & 0.000 \\
+    60-69 & 4321 & 2601 & 9 \\
+    & 0.094 & 0.056 & 0.000 \\
+    50-59 & 3463 & 2151 & 9 \\
+    & 0.075 & 0.047 & 0.000 \\
+    90+ & 2085 & 2816 & 13 \\
+    & 0.045 & 0.061 & 0.000 \\
+    40-49 & 1846 & 1110 & 5 \\
+    & 0.040 & 0.024 & 0.000 \\
+    30-39 & 784 & 756 & 4 \\
+    & 0.017 & 0.016 & 0.000 \\
+    \hline
+    \end{tabular}
+\end{table}
+
+## The Null Kaplan-Meier Curves
+
+```{r fig.height = 3.5, fig.align = 'center', warning = FALSE, message = FALSE, fig.cap = "The null survival curves, whereby (A) is based on complete cases only, whilst (B) is based on all plausible records after multiple imputation."}
+KaplanMeier()
+```
+
+## The Cox Model Assumptions
+
+**The observations are independent**: This assumption requires a set of observations wherein the patients are not related or associated.
+If there are clusters of relations or associates, the clusters should be, must be, indicated as-such.
+By virtue of the > study's design, it is quite possible that the observations are not independent - families, friends, and colleagues living in the same area will probably be admitted to the same health centre.
+Alas, the data set does not have cluster indicators, hence it is impossible to definitively test and state the independence, or otherwise, of the observations.
+
+**Independent Censoring**:  There is neither evidence to suggest nor rule-out <span style="color: brown;">outcomes due to systematic drop-outs, induced by study design flaws</span>.
+What are the prognosis of patients released to palliative care or transferred?
+Is the decision to transfer or release to palliative care influenced or dictated by *lower* survival probability?
+If the answer to the latter question is yes - it will be in breach of independent censoring.
+
+> **Proportional Hazards**
+>
+> *The curves.*
+>
 
@@
 
-# Equations
+# Discussion
 
@@
 
-# Inserting R figures
+# Conclusion
 
@@
 
-## Figures spanning two-columns
-
 Figures can span two columns be setting `fig.env="figure*"`.
@@
 Reference to second figure: Fig. \ref{fig:fig2}
-
-# Tables
 
@@
 
-# Cross-referencing sections
+\appendix
 
-You can cross-reference sections and subsections as follows: Section \ref{literature-citations} and Section \ref{a-subsection}.
+\section{Appendices}
 
-***Note:*** the last section in the document will be used as the section title for the bibliography.
+\subsection{Exploratory Analysis}
 
-For more portable and flexible referencing of sections, equations, figures and tables, use [`bookdown::pdf_document2`](https://github.com/rstudio/bookdown) with YAML header option `base_format: rticles::oup_article`.
+Graphs
 
-# Appendices {-}
+\subsection{Missing Data}
 
-::: {.appendices latex=true}
-
-# Section title of first appendix
-
-blabla
-
-## Subsection title of first appendix
-
-and so on....
-:::
+Clusters, patterns, associations.
 

### 2022-01-18 R/evaluation/Characteristics.R
Add R/evaluation/Characteristics.R

@@ -1,0 +1,40 @@
+# Title     : Characteristics.R
+# Objective : Characteristics
+# Created by: greyhypotheses
+# Created on: 16/01/2022
+
+source(file = 'R/functions/StudyData.R')
+
+Characteristics <- function () {
+
+  data <- StudyData()
+
+  # Mortality
+  deaths <- table(data$deceased, useNA = 'always')
+  rbind(deaths, round(prop.table(deaths), digits = 3))
+
+  # Demographics
+  demographics <- table(data$sex, data$age_group, useNA = 'always')
+  t(demographics)
+  t(round(prop.table(demographics), digits = 3))
+  brief <- data %>%
+    select(sex, age_group) %>%
+    group_by(sex, age_group) %>%
+    summarise(N = n()) %>%
+    pivot_wider(id_cols = age_group, names_from = sex, values_from = N) %>%
+    data.frame()
+  brief
+
+  # Diseases
+  diseases <- c('asthma', 'liver_mild', 'renal', 'pulmonary', 'neurological', 'liver_mod_severe', 'malignant_neoplasm')
+  frequencies <- function (field) {
+    N <- table(data[, field], useNA = 'always')
+    fractions <- round(prop.table(N), digits = 2)
+    frame <- data.frame(rbind(N, fractions))
+    row.names(frame) <- c(field, 'fractions')
+    return(frame)
+  }
+  details <- lapply(X = diseases, FUN = frequencies)
+  dplyr::bind_rows(details)
+
+}

### 2022-01-18 R/evaluation/Characteristics.R
Update R/evaluation/Characteristics.R

@@ -14,27 +14,21 @@ Characteristics <- function () {
   rbind(deaths, round(prop.table(deaths), digits = 3))
 
   # Demographics
-  demographics <- table(data$sex, data$age_group, useNA = 'always')
-  t(demographics)
-  t(round(prop.table(demographics), digits = 3))
-  brief <- data %>%
+  demographics <- data %>%
     select(sex, age_group) %>%
     group_by(sex, age_group) %>%
     summarise(N = n()) %>%
     pivot_wider(id_cols = age_group, names_from = sex, values_from = N) %>%
     data.frame()
-  brief
+  demographics
 
   # Diseases
-  diseases <- c('asthma', 'liver_mild', 'renal', 'pulmonary', 'neurological', 'liver_mod_severe', 'malignant_neoplasm')
-  frequencies <- function (field) {
-    N <- table(data[, field], useNA = 'always')
-    fractions <- round(prop.table(N), digits = 2)
-    frame <- data.frame(rbind(N, fractions))
-    row.names(frame) <- c(field, 'fractions')
-    return(frame)
-  }
-  details <- lapply(X = diseases, FUN = frequencies)
-  dplyr::bind_rows(details)
+  diseases <- data %>%
+    select(asthma, liver_mild, renal, pulmonary, neurological, liver_mod_severe, malignant_neoplasm) %>%
+    gather(key = 'disease', value = 'Class') %>%
+    group_by(disease, Class) %>%
+    summarise(N = n(), .groups = 'drop') %>%
+    pivot_wider(id_cols = 'disease', names_from = 'Class', values_from = 'N') %>%
+    data.frame()
 
 }

### 2022-01-18 R/background/Numbers.R
Rename R/background/Characteristics.R -> R/background/Numbers.R

@@ -3,7 +3,7 @@
 # Created by: greyhypotheses
 # Created on: 16/01/2022
 
-source(file = 'R/functions/StudyData.R')
+source(file = '../functions/StudyData.R')
 
 Characteristics <- function () {
 

### 2022-01-18 R/background/Numbers.R
Update R/background/Numbers.R

@@ -1,4 +1,4 @@
-# Title     : Characteristics.R
+# Title     : Numbers.R
 # Objective : Characteristics
 # Created by: greyhypotheses
 # Created on: 16/01/2022

### 2022-01-18 manuscript.Rmd
Update manuscript.Rmd

@@
 author: 36112985
-abstract:
-  - "**Background:**"
-  - "**Objective:**  To Investigate, and identify, the factors that may influence patient survival of coronavirus disease 19."
-  - "**Methods**"
-  - "**Motivation:** You can also have some paragraphs start with bold face."
 bibliography: manuscript/bibliography.bib
@@
 sys.source(file = 'R/functions/StudyData.R', envir = knitr::knit_global())
+
+sys.source(file = 'R/background/Numbers.R', envir = knitr::knit_global())
 ```
+
+```{r}
+data <- StudyData()
+```
+
+> # ABSTRACT
+>
+> **Objective** *To investigate, amongst a set of probable prognostic factors, which factors may influence coronavirus 19 disease survival probabilities.*
+>
+> **Design** *Prospective cohort study; ongoing.*
+>
+> **Setting** *The International Severe Acute Respiratory and Emerging Infections Consortium (ISARIC) World Health Organization (WHO) Clinical Characterisation Protocol UK (CCP-UK) study, which enrols in-patients at at-least 260 England, Scotland, and Wales hospitals; the ISARIC Coronavirus Clinical Characterisation Consortium (ISARIC 4C) conducts the study.
+The investigation's patients are patients enrolled in a participating England hospital between 10 February 2020 and 5 July 2020.*
+>
+> **Participants** *Each study enrollee has a confirmed SARS-CoV-2 infection, or a high infection likelihood.  The investigation's patients are members of age groups *30 - 39*, *40 - 49*, and higher.*
+>
+> **Outcome** Patient survival.
+>
+> **Results** Inconclusive.
+>
+> **Conclusions**  Insufficient spectrum of factors.
 
@@
 
-During the past 2years, infections due to severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2) have led to a high number of coronavirus 19 disease deaths and in-hospital care.
-With time, case fatality rates, and in-hospital survival rates, have improved [...] as our understanding of coronavirus 19 disease pathogenesis has developed, and hence prevention, mitigation, and treatment measures have improved.
+During the past 2 years, infections due to severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2) have led to a high number of coronavirus 19 disease deaths and in-hospital care.
+With time, case fatality rates, and in-hospital survival rates, have improved [...] as our understanding of coronavirus 19 disease pathogenesis has developed.
+Consequently,  disease prevention, mitigation, and treatment measures have improved.
+
 Effective disease treatment measures depend on our understanding of the disease's prognosis factors [@riley2013], the focus of prognostic factors research.
@@
 The study's hypothesis is that *one or more of the measures has a significant impact on each patient's survival probability*.
+The ISARIC 4C study outcomes are: discharged alive, remains in hospital, death, transferred, and palliative discharge.
+Herein, the overarching outcomes of interest are: dead, alive.
+The predictors, factors, under investigation
 
@@
 
-## Outcomes
+## Exploratory Data Analysis
 
-The ISARIC 4C study outcomes are: discharged alive, remains in hospital, death, transferred, and palliative discharge. Herein, the overarching outcomes of interest are: dead, alive.
+* Relationships
+* Correlations
 
@@
 
-## Independent Variables
+## Missing Data
 
-The predictors, factors, under investigation
+Most of the data's variables have missing values.
+Prior to deciding how to address missing values, it is important to understand the missing values patterns [@steyerberg2010].
+Rubin [@rubin1976, @steyerberg2010, @little2019] outlines three fundamental missing data mechanisms
 
-\vspace{20pt}
+* missing completely at random (MCAR) / administrative errors, accidents
+* missing at random (MAR) / missing data associated with known patient characteristics [independent variables], or the outcome
+* missing not at random (MNAR) / missing data associated with missing values of the factor/predictor in question or with unobserved predictors
+
+If the missing values of a data set in question are *missing completely at random* then complete case analysis will suffice because the complete case excerpt is akin to a random sample from a complete population.
+If MCAR does not hold, e.g., data is *missing at random*, then the complete case excerpt is not representative of the underlying population, therefore population inference is not possible via complete case analysis.
+[@steyerberg2010]
+
+Herein, missing data mechanisms analysis is via logistic regression.
+Almost all the independent variables have missing values patterns that violate MCAR; addressed via multiple imputation in-line with the advice and protocols of [@sterne2009] & [@steyerberg2010].
 
@@
 
+## Null Survival Curve
+
+* Complete Case
+* Imputation Case
+
+## Uni-variate & Multivariate Analysis
+
+* The table ... Cox
+
+## Adjusted Analysis via Restricted Mean Survival Time
+
+The proportional hazard assumption does not hold ...
+
+# Discussion
+
 ## The Cox Model Assumptions
@@
 
-> **Proportional Hazards**
->
-> *The curves.*
->
+**Proportional Hazards**: The curves.
 
-By default, citations are handled by `natbib` using a numeric citation format. To use name-date citations, sets `namedate: TRUE` in the YAML header.
+## Missing Data
 
-Here are two sample references:
+Within a missing values setting, there are 2 options w.r.t. predictor effect analysis [@steyerberg2010]
 
-* **author (year) example:** @horvath2018dna showed some really cool things. Only seems to work properly if `namedate: TRUE`. 
-* **(author year) example:** This is a well known result [@ji20123d].
+* complete case analysis for the uni-variate analysis, and complete case predictors & imputed confounding variables for adjusted analysis.
+* uni-variate and adjusted analysis after imputation of all missing values.
 
-The bibliography will appear at the end of the document.
+The investigation opted for the latter, especially because ...
 
-Though not normally available in the OUP LaTeX format, [CSL style files](https://www.zotero.org/styles) can also be used with the Rmarkdown adaptation by setting in the YAML header `citation_package: "default"` and defining the `csl` element to be the path towards the style file.
+## Bias
 
-# Discussion
+Possible bias points
 
-An equation without a label for cross-referencing:
+* By virtue of the study's design pre-study--admission differences between hospitals is possible due to the ambiguity of the study's patient recruitment terms, and this might lead to post-study differences.
+In a nutshell, the study is susceptible to selection bias.
+[Trochim]
 
-$$
-E=mc^2
-$$
-
-An inline equation: $y=ax+b$
-
-An equation with a label for cross-referencing:
-
-\begin{equation}\label{eq:eq1}
-\int^{r_2}_0 F(r,\varphi){\rm d}r\,{\rm d}\varphi = 1
-\end{equation}
-
-This equation can be referenced as follows: Eq. \ref{eq:eq1}
+* Missing data [@sterne2009]
 
@@
 
-The code below creates a figure.
-
-```{r fig1, fig.cap="This is the first figure."}
-plot(1:10,main="Some data",xlab="Distance (cm)",
-     ylab="Time (hours)")
-```
-
-You can reference this figure as follows: Fig. \ref{fig:fig1}.
-
-Figures can span two columns be setting `fig.env="figure*"`.
-
-```{r fig2, fig.cap="This is a wide figure.", fig.env="figure*"}
-plot(1:5,pch=19,main="Some wide data",
-     xlab="Distance (cm)",ylab="Time (hours)")
-```
-
-Reference to second figure: Fig. \ref{fig:fig2}
-
-## Generate a table using `xtable`
-
-```{r xtabletable, results="asis", echo=TRUE}
-df <- data.frame(ID=1:3,code=letters[1:3])
-
-# Creates tables that follow OUP guidelines 
-# using xtable
-
-print(xtable::xtable(df,caption="This is a xtable table.", label="tab:tab1"), comment=FALSE,caption.placement="top")
-```
-
-You can reference this table as follows: Table \ref{tab:tab1}.
-
-## Generate a table using `kable`
-
-```{r kabletable, echo=TRUE}
-df <- data.frame(ID=1:3,code=letters[1:3])
-
-# kable can alse be used for creating tables
-knitr::kable(df,caption="This is a kable table.",
-             booktabs=TRUE,label="tab2")
-```
-
-You can reference this table as follows: Table \ref{tab:tab2}.
-
-## Table spanning two columns
-
-Tables can span two columns be setting `table.envir = "table*"` in `knitr::kable`.
-
-```{r widetable, echo=TRUE}
-df <- data.frame(ID=1:3,code1=letters[1:3],
-                code2=letters[4:6],
-                code3=letters[7:9],
-                code4=letters[10:12],
-                code5=letters[13:15])
-
-# kable can alse be used for creating tables
-knitr::kable(df,caption="This is a wide kable table.",
-             #format="latex",
-             table.envir="table*",
-             booktabs=TRUE,label="tab3")
-```
+The analysis of missing values hints at flaws in data collection.
 
@@
 
+```{r}
+
+```
+
 Graphs
@@
 
+<!--- # References {#references .unnumbered} --->
 # References

### 2022-01-18 paper.Rmd
Update paper.Rmd

@@
     corresponding: 36112985@lancaster.ac.uk
-    affiliation: 
-      - Lancaster Medical School
 address:
@@
 abstract: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
-
-author_summary: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+  **OBJECTIVE** To investigate, amongst a set of probable prognostic factors, which factors may influence coronavirus 19 disease survival probabilities.
+**DESIGN** Prospective cohort study; ongoing.
+**SETTING** The International Severe Acute Respiratory and Emerging Infections Consortium (ISARIC) World Health Organization (WHO) Clinical Characterisation Protocol UK (CCP-UK) study, which enrols in-patients at at-least 260 Englsnd, Scotland, and Wales hospitals; the ISARIC Coronavirus Clinical Characterisation Consortium (ISARIC 4C) conducts the study.
+The investigation's patients are patients enrolled in a participating England hospital between 10 February 2020 and 5 July 2020.
+PARTICIPANTS  Each study enrollee has a confirmed SARS-CoV-2 infection, or a high infection likelihood.
+The investigation's patients are members of age groups *30 - 39*, *40 - 49*, and higher.
+**OUTCOME** Patient survival.
+**RESULTS** Inconclusive.
+**CONCLUSIONS**  Insufficient spectrum of factors.
 
@@
 
-## Null Survival Curve
-
-* Complete Case
-* Imputation Case
-
-## Uni-variate Analysis
-
-* The table ... Cox
-
-## Adjusted Analysis via Restricted Mean Survival Time
-
-The proportional hazard assumption does not hold ...
-
-# Results
-
-# Discussion
-
-Within a missing values setting, there are 2 options w.r.t. predictor effect analysis [@steyerberg2010]
-
-* complete case analysis for the uni-variate analysis, and complete case predictors & imputed confounding variables for adjusted analysis.
-* uni-variate and adjusted analysis after imputation of all missing values.
-
-The investigation opted for the latter, especially because ...
-
-## Bias
-
-Possible bias points
-
-* By virtue of the study's design pre-study--admission differences between hospitals is possible due to the ambiguity of the study's patient recruitment terms, and this might lead to post-study differences.
-In a nutshell, the study is susceptible to selection bias.
-[Trochim]
-
-* Missing data [@sterne2009]
-
-# Conclusion
-
-The analysis of missing values hints at flaws in data collection.
-
 # References {#references .unnumbered}

### 2022-01-18 R/background/DiseaseGraphNumbers.R
Update R/background/DiseaseGraphNumbers.R

@@ -1,10 +1,10 @@
-# Title     : DiseaseNumbers.R
+# Title     : DiseaseGraphNumbers.R
 # Objective : yes/no numbers of a disease by age group & sex
 # Created by: greyhypotheses
 # Created on: 04/01/2022
 
 
-DiseaseNumbers <- function (field, original) {
+DiseaseGraphNumbers <- function (field, original) {
 
   excerpt <- original[, c('age_group', 'sex', field)]
 

### 2022-01-18 R/background/DiseaseGraphQuotients.R
Update R/background/DiseaseGraphQuotients.R

@@ -1,9 +1,9 @@
-# Title     : DiseaseQuotients.R
+# Title     : DiseaseGraphQuotients.R
 # Objective : yes/no quotients of a disease by age group & sex
 # Created by: greyhypotheses
 # Created on: 04/01/2022
 
-DiseaseQuotients <- function (field, original) {
+DiseaseGraphQuotients <- function (field, original) {
 
   excerpt <- original[, c('age_group', 'sex', field)]
 

### 2022-01-18 R/background/Numbers.R
Update R/background/Numbers.R

@@ -1,28 +1,45 @@
 # Title     : Numbers.R
-# Objective : Characteristics
+# Objective : Numbers
 # Created by: greyhypotheses
 # Created on: 16/01/2022
 
-source(file = '../functions/StudyData.R')
 
-Characteristics <- function () {
+#' Deaths
+DeceasedNumbers <- function (data) {
 
-  data <- StudyData()
-
-  # Mortality
   deaths <- table(data$deceased, useNA = 'always')
-  rbind(deaths, round(prop.table(deaths), digits = 3))
+  return(data.frame(yes = deaths['1'],
+                    no = deaths['0'],
+                    unknown = deaths[length(deaths)],
+                    row.names = 'deceased'))
+}
 
-  # Demographics
+#' Demographics
+DemographicsNumbers <- function (data) {
+
   demographics <- data %>%
     select(sex, age_group) %>%
     group_by(sex, age_group) %>%
-    summarise(N = n()) %>%
+    summarise(N = n(), .groups = 'drop') %>%
     pivot_wider(id_cols = age_group, names_from = sex, values_from = N) %>%
     data.frame()
-  demographics
 
-  # Diseases
+  demographics$Total <- select(demographics, !age_group) %>% rowSums()
+  names(demographics) <- c('Age Group', 'Male', 'Female', 'Unknown', 'Total')
+
+  return(demographics)
+
+}
+
+#' Diseases
+DiseaseNumbers <- function (data) {
+
+  disease <- c('asthma', 'liver_mild', 'liver_mod_severe',
+            'malignant_neoplasm', 'neurological', 'pulmonary', 'renal')
+  name <- c('Asthma', 'Mild Liver Disease', 'Moderate, Severe Liver Disease',
+            'Malignant Neoplasm', 'Neurological Disorder', 'Pulmonary Disease', 'Renal Disease')
+  labels <- data.frame(disease = disease, Comorbidity = name)
+
   diseases <- data %>%
     select(asthma, liver_mild, renal, pulmonary, neurological, liver_mod_severe, malignant_neoplasm) %>%
     gather(key = 'disease', value = 'Class') %>%
@@ -30,5 +47,15 @@ Characteristics <- function () {
     summarise(N = n(), .groups = 'drop') %>%
     pivot_wider(id_cols = 'disease', names_from = 'Class', values_from = 'N') %>%
     data.frame()
[…]

### 2022-01-18 R/background/interface.R
Update R/background/interface.R

@@ -8,8 +8,8 @@
 source(file = 'R/functions/StudyData.R')
 
 source(file = 'R/background/AgeGroupSexEvent.R')
-source(file = 'R/background/DiseaseNumbers.R')
-source(file = 'R/background/DiseaseQuotients.R')
+source(file = 'R/background/DiseaseGraphNumbers.R')
+source(file = 'R/background/DiseaseGraphQuotients.R')
 source(file = 'R/background/CorrelationOfPredictors.R')
 
 source(file = 'R/background/QuotientsDistributions.R')
@@ -35,12 +35,12 @@ AgeGroupSexEvent(original = data)
 #' ... Although the absolute numbers differ between the sexes, the pattern of
 #'     sufferers-by-age-group is similar between the sexes.
 #'
-DiseaseNumbers(field = 'pulmonary', original = data)
+DiseaseGraphNumbers(field = 'pulmonary', original = data)
 
 
 #' Illustration: Per comorbidity, the yes/no ...
 #'
-DiseaseQuotients(field = 'pulmonary', original = data)
+DiseaseGraphQuotients(field = 'pulmonary', original = data)
 
 
 #' Correlation

### 2022-01-19 R/modelling/Multivariate.R
Update R/modelling/Multivariate.R

@@ -36,7 +36,7 @@ Multivariate <- function(blob) {
   frame[indices, 'proportionality'] <- evidence[row.names(evidence) == 'age_group',]
 
   # field names
-  names(frame) <- c('Factor', 'Coefficient', 'HR', 'HRLCI', 'HRUCI', 'p_value', 'proportionality')
+  names(frame) <- c('label', 'Coefficient', 'HR', 'HRLCI', 'HRUCI', 'p_value', 'proportionality')
 
   return(list(diagnostics = frame, model = model))
 

### 2022-01-19 R/modelling/Univariate.R
Update R/modelling/Univariate.R

@@ -44,6 +44,9 @@ Univariate <- function (blob) {
   }  
   est <- lapply(X = models, FUN = function (model){estimates(model = model)})
   diagnostics <- dplyr::bind_rows(est)
+
+
+
   
   return(list(diagnostics = diagnostics, models = models))
 }

### 2022-01-19 R/modelling/interface.R
Update R/modelling/interface.R

@@ -19,28 +19,23 @@ source(file = 'R/evaluation/SurvivalCurve.R')
 dataframes <- ExtensiveStudyData(upload = TRUE)
 data <- dataframes$data
 training_ <- dataframes$training_
-testing_ <- dataframes$testing_
 data_ <- dataframes$data_
-
-even <- data[complete.cases(data),]
-
-source(file = 'R/modelling/Univariate.R')
-analysis <- Univariate(blob = training_)
-analysis <- Univariate(blob = data_)
-analysis <- Univariate(blob = even)
 
 
 source(file = 'R/modelling/Multivariate.R')
 analysis <- Multivariate(blob = training_)
 analysis <- Multivariate(blob = data_)
-analysis <- Multivariate(blob = even)
+analysis <- Multivariate(blob = data[complete.cases(data), ])
+
+diagnostics <- analysis$diagnostics
+model <- analysis$model
 
 
 #' Training
 #'
 #' The Cox model is inappropriate, proportionality assumption does not hold.
 #'
-#' For time depedent cases:
+#' For time dependent cases:
 #' https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Time-dependent_covariate
 #' https://glmnet.stanford.edu/articles/Coxnet.html#cox-models-for-start-stop-data-1
 #' https://www.rdocumentation.org/packages/glmnet/versions/4.1-3/topics/glmnet
@@ -51,8 +46,8 @@ analysis <- Multivariate(blob = even)
 
 SurvivalCurve(data = training_, caption = 'Training Data (multiply imputed)')
 ViolationsKaplan(data = training_)
-ViolationsTime(model = unboosted)
-ViolationsProportionalityTable(model = unboosted)
+ViolationsTime(model = model)
+ViolationsProportionalityTable(model = model)
 
 
 

### 2022-01-19 manuscript.Rmd
Update manuscript.Rmd

@@
   - \usepackage{caption}
+  - \usepackage{booktabs}
   - \captionsetup[figure]{font={normalsize, color=gray}, width=.8\linewidth}
@@
 library(patchwork)
+library(kableExtra)
 ```
@@
 sys.source(file = 'R/evaluation/KaplanMeier.R', envir = knitr::knit_global())
-sys.source(file = 'R/functions/StudyData.R', envir = knitr::knit_global())
-
+sys.source(file = 'R/functions/ExtensiveStudyData.R', envir = knitr::knit_global())
+sys.source(file = 'R/modelling/UnivariateAnalysis.R', envir = knitr::knit_global())
+sys.source(file = 'R/modelling/MultivariateAnalysis.R', envir = knitr::knit_global())
 sys.source(file = 'R/background/Numbers.R', envir = knitr::knit_global())
@@
 ```{r}
-data <- StudyData()
+dataframes <- ExtensiveStudyData(upload = TRUE)
+data <- dataframes$data
+training_ <- dataframes$training_
+data_ <- dataframes$data_
 ```
@@
 \begin{table}[h!]
+    \scriptsize
     \centering
@@
 
+Correlation graph
+
+## Missing Data
+
 ## The Null Kaplan-Meier Curves
@@
 
-## Null Survival Curve
+## Uni-variate & Multivariate Analysis: Training
 
-* Complete Case
-* Imputation Case
+The results of Table \@ref(tab:univariateTraining) summarise the separate effects of each potential prognostic factor on survival probability; investigated via Cox Proportional Hazard uni-variate analysis $\rightarrow$ coefficient: regression coefficients, hazard ratio & confidence interval: effect size, $p_{value}$: statistical significance in relation to survival.
 
-## Uni-variate & Multivariate Analysis
+In terms of age groups, the reference age group is *80 – 89*, and their $p_{value}$ scores suggests that **(a)** each age group, except *90+*, has a lower risk of death compared to age group *80 – 89*, and **(b)** the risk of death increases with age.
+The sex $p_{value}$ suggests that females have a lower risk of death compared to males.
+The $p_{value}$ scores for patients with an asthma, renal disease, pulmonary disease, or malignant neoplasm, diagnosis suggests that these patients have a higher risk of death compared to their respective disease/disorder free counterparts.
 
-* The table ... Cox
+Alas, the proportionality tests invalidate most of the preceding observations; age group and malignant neoplasm have a score test value below $0.05$, the renal disease score test value is approximately $0.05$, and neurological disorder score is barely greater than $0.05$.
+The multivariate analysis confirms the trend.
+
+\footnotesize
+```{r univariateTraining}
+diagnostics <- UnivariateAnalysisTraining()
+
+diagnostics %>%
+  dplyr::mutate_all(linebreak) %>%
+  kableExtra::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    digits = 3,
+    escape = FALSE,
+    col.names = kableExtra::linebreak(c('', 'coefficient', 'hazard\nratio', 'lower',
+                                        'upper', '$p_{value}$', '${U}$' )),
+    caption = 'Univariate Analysis', position = '!ht') %>%
+    kableExtra::kable_styling(position = 'center') %>%
+    kableExtra::add_header_above(c('', '', '', 'hazard ratio\nCI' = 2, '', '')) %>%
+    kableExtra::add_footnote(c('CI: Confidence Interval',
+                               'U: Proportionality score test'), notation = "number") %>%
+    kableExtra::column_spec(column = 1, latex_column_spec = "p{0.20\\\\linewidth}l") %>%
+    kableExtra::column_spec(column = 2, latex_column_spec = "r")
+```
+\normalsize
+
+\footnotesize
+```{r multivariateTraining}
+diagnostics <- MultivariateAnalysisTraining()
+
+diagnostics %>%
+  dplyr::mutate_all(linebreak) %>%
+  kableExtra::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    digits = 3,
+    escape = FALSE,
+    col.names = kableExtra::linebreak(c('', 'coefficient', 'hazard\nratio', 'lower',
+                                        'upper', '$p_{value}$', '${U}$' )),
+    caption = 'Multivariate Analysis', position = '!ht') %>%
+  kableExtra::kable_styling(position = 'center') %>%
+  kableExtra::add_header_above(c('', '', '', 'hazard ratio\nCI' = 2, '', '')) %>%
+  kableExtra::add_footnote(c('CI: Confidence Interval',
+                             'U: Proportionality score test'), notation = "number")  %>%
+  kableExtra::column_spec(column = 1, latex_column_spec = "p{0.20\\\\linewidth}l") %>%
+  kableExtra::column_spec(column = 2, latex_column_spec = "r")
+```
+\normalsize
+
+## Uni-variate & Multivariate Analysis: All
+
+Due to the failure of the Cox Proportional Hazard Modelling & Analysis strategy, ...
 
@@
 
+\pagebreak
+
 \appendix
@@
 ```{r}
-
+blob <- DiseaseNumbers(data = data)
+knitr::kable(blob, caption = 'The disease frequencies', format = 'latex', booktabs = TRUE)
 ```
@@
 
+\pagebreak
 <!--- # References {#references .unnumbered} --->

### 2022-01-19 R/main.R
Update R/main.R

@@ -24,6 +24,10 @@ rm(dataframes)
 str(data)
 
 
+#' Basic characteristics
+c(starting = min(data$admission_date), ending = max(data$admission_date))
+table(data$age_group)
+
 
 #' Is()
 #'

### 2022-01-19 R/modelling/MultivariateAnalysis.R
Add R/modelling/MultivariateAnalysis.R

@@ -1,0 +1,57 @@
+# Title     : MultivariateAnalysis.R
+# Objective : Multivariate Analysis
+# Created by: greyhypotheses
+# Created on: 19/01/2022
+
+
+source(file = 'R/modelling/Multivariate.R')
+source(file = 'R/functions/ExtensiveStudyData.R')
+
+
+MultivariateAnalysisTraining <- function () {
+
+  dataframes <- ExtensiveStudyData(upload = TRUE)
+  training_ <- dataframes$training_
+
+  analysis <- Multivariate(blob = training_)
+  diagnostics <- analysis$diagnostics
+  diagnostics <- diagnostics %>% arrange(label)
+
+  # names, labels, indices
+  settings <- data.frame(label = c('age_group70-79', 'age_group60-69', 'age_group50-59', 'age_group90+', 'age_group40-49',
+                                   'age_group30-39', 'sexFemale', 'asthmayes', 'liver_mildyes', 'renalyes', 'pulmonaryyes',
+                                   'neurologicalyes', 'liver_mod_severeyes', 'malignant_neoplasmyes'),
+                         name = c('Age Group 70 - 79', 'Age Group 60 - 69', 'Age Group 50 - 59', 'Age Group 90+',
+                                  'Age Group 40 - 49', 'Age Group 30 - 39', 'Sex Female', 'Asthma Yes',
+                                  'Liver Disease, (Mild) Yes', 'Renal Disease Yes', 'Pulmonary Disease Yes',
+                                  'Neurological Disorder Yes', 'Liver Disease, (Moderate, severe) Yes', 'Malignant Neoplasm Yes'))
+
+  diagnostics <- dplyr::left_join(x = diagnostics, y = settings, by = 'label')
+  diagnostics <- diagnostics %>%
+    select(name, Coefficient, HR, HRLCI, HRUCI, p_value, proportionality)
+
+  return(diagnostics)
+
+
+}
+
+
+MultivariateAnalysisData <- function () {
+
+  dataframes <- ExtensiveStudyData(upload = TRUE)
+  data_ <- dataframes$data_
+
+  analysis <- Multivariate(blob = data_)
+  diagnostics <- analysis$diagnostics
+
+}
+
+
+MultivariateAnalysisCC <- function () {
+
+  dataframes <- ExtensiveStudyData(upload = TRUE)
+  data <- dataframes$data
+
+  C <- data[complete.cases(data),]
+  analysis <- Multivariate(blob = C)
+}

### 2022-01-19 R/modelling/UnivariateAnalysis.R
Add R/modelling/UnivariateAnalysis.R

@@ -1,0 +1,74 @@
+# Title     : UnivariateAnalysis.R
+# Objective : Analysis
+# Created by: greyhypotheses
+# Created on: 19/01/2022
+
+
+source(file = 'R/modelling/Univariate.R')
+source(file = 'R/functions/ExtensiveStudyData.R')
+
+UnivariateAnalysisTraining <- function () {
+
+  dataframes <- ExtensiveStudyData(upload = TRUE)
+  training_ <- dataframes$training_
+
+  analysis <- Univariate(blob = training_)
+  diagnostics <- analysis$diagnostics
+
+  # names, labels, indices
+  diagnostics$label <- row.names(diagnostics)
+  row.names(diagnostics) <- NULL
+  diagnostics <- diagnostics %>% arrange(label)
+
+  settings <- data.frame(label = c('age_group70-79', 'age_group60-69', 'age_group50-59', 'age_group90+', 'age_group40-49',
+                                   'age_group30-39', 'sexFemale', 'asthmayes', 'liver_mildyes', 'renalyes', 'pulmonaryyes',
+                                   'neurologicalyes', 'liver_mod_severeyes', 'malignant_neoplasmyes'),
+                         name = c('Age Group 70 - 79', 'Age Group 60 - 69', 'Age Group 50 - 59', 'Age Group 90+',
+                                  'Age Group 40 - 49', 'Age Group 30 - 39', 'Sex Female', 'Asthma Yes',
+                                  'Liver Disease, (Mild) Yes', 'Renal Disease Yes', 'Pulmonary Disease Yes',
+                                  'Neurological Disorder Yes', 'Liver Disease, (Moderate, severe) Yes', 'Malignant Neoplasm Yes'))
+
+  diagnostics <- dplyr::left_join(x = diagnostics, y = settings, by = 'label')
+  diagnostics <- diagnostics %>%
+    select(name, Coefficient, HR, HRLCI, HRUCI, p_value, proportionality)
+
+  return(diagnostics)
+
+
+}
+
+
+UnivariateAnalysisData <- function () {
+
+  dataframes <- ExtensiveStudyData(upload = TRUE)
+  data_ <- dataframes$data_
+
+  analysis <- Univariate(blob = data_)
+  diagnostics <- analysis$diagnostics
+
+
+
+}
+
+
+UnivariateAnalysisCC <- function () {
+
[…]

### 2022-01-19 R/modelling/UnivariateAnalysis.R
Update R/modelling/UnivariateAnalysis.R

@@ -1,5 +1,5 @@
 # Title     : UnivariateAnalysis.R
-# Objective : Analysis
+# Objective : Univariate Analysis
 # Created by: greyhypotheses
 # Created on: 19/01/2022
 

### 2022-01-19 R/restricted/Adjusted.R
Add R/restricted/Adjusted.R

@@ -1,0 +1,39 @@
+# Title     : Adjusted.R
+# Objective : Adjusted
+# Created by: greyhypotheses
+# Created on: 19/01/2022
+
+
+Adjusted <- function (excerpt, time, status, tau, keys) {
+
+  K <- c(keys, 'age_group_median')
+
+
+  # adjusted RMST function
+  adjusted <- function (key) {
+
+    # modelling
+    arm <- excerpt[, key]
+    covariates <- excerpt[, K[K != key]]
+    obj <- rmst2(time = time, status = status, arm = arm, tau = tau, covariates = covariates)
+
+    # results
+    reference <- obj
+    estimates <- reference$adjusted.result
+    contrast <- estimates[row.names(estimates) == 'RMST (arm=1)-(arm=0)', ]
+
+    # summary
+    return(data.frame(est_ = contrast['Est.'], lower_ = contrast['lower .95'],
+                      upper_ = contrast['upper .95'], p_ = contrast['p'], row.names = key))
+
+  }
+
+
+  # calculate per factor
+  calculations <- dplyr::bind_rows(lapply(X = keys, FUN = function (x){adjusted(x)}))
+
+
+  return(calculations)
+
+
+}

### 2022-01-19 R/restricted/AgeGroupMedians.R
Add R/restricted/AgeGroupMedians.R

@@ -1,0 +1,25 @@
+# Title     : AgeGroupMedians.R
+# Objective : Age Group Medians
+# Created by: greyhypotheses
+# Created on: 19/01/2022
+
+
+AgeGroupMedians <- function (blob) {
+
+  # data frame of age groups
+  estimates <- data.frame(lower = seq(30, 90, 10), upper = seq(39, 99, 10))
+  estimates$age_group <- paste(estimates$lower, estimates$upper, sep = '-')
+  estimates$age_group <- str_replace(string = estimates$age_group, pattern = '90-99', replacement = '90+')
+
+  # their medians
+  estimates <- estimates %>%
+    mutate(age_group_median = rowMedians(as.matrix(estimates[, c('lower', 'upper')])))
+
+  # associate each age group element of blob with its mean
+  blob <- dplyr::left_join(x = blob,
+                           y = estimates[, c('age_group', 'age_group_median')],
+                           by = 'age_group')
+
+  return(blob)
+  
+}

### 2022-01-19 R/restricted/Unadjusted.R
Add R/restricted/Unadjusted.R

@@ -1,0 +1,34 @@
+# Title     : Unadjusted.R
+# Objective : Unadjusted
+# Created by: greyhypotheses
+# Created on: 19/01/2022
+
+
+Unadjusted <- function (excerpt, time, status, tau, keys) {
+
+  # unadjusted RMST function
+  unadjusted <- function (key) {
+
+    # modelling
+    arm <- excerpt[, key]
+    obj <- survRM2::rmst2(time = time, status = status, arm = arm, tau = tau)
+
+    # results
+    reference <- obj
+    estimates <- reference$unadjusted.result
+    contrast <- estimates[row.names(estimates) == 'RMST (arm=1)-(arm=0)', ]
+
+    # summary
+    return(data.frame(est = contrast['Est.'], lower = contrast['lower .95'],
+                      upper = contrast['upper .95'], p = contrast['p'], row.names = key))
+
+  }
+
+  # calculate per factor
+  calculations <- dplyr::bind_rows(lapply(X = keys, FUN = function (x){unadjusted(x)}))
+
+
+  return(calculations)
+
+
+}

### 2022-01-19 R/restricted/interface.R
Add R/restricted/interface.R

@@ -1,0 +1,81 @@
+# Title     : interface.R
+# Objective : Interface for restricted mean surivival time analysis
+# Created by: greyhypotheses
+# Created on: 06/01/2022
+
+
+
+#' Programs
+#'
+source(file = 'R/functions/ExtensiveStudyData.R')
+source(file = 'R/restricted/AgeGroupMedians.R')
+source(file = 'R/restricted/Unadjusted.R')
+source(file = 'R/restricted/Adjusted.R')
+
+
+# data
+dataframes <- ExtensiveStudyData(upload = TRUE)
+data_ <- dataframes$data_
+
+
+# for univariate RMST
+keys <- c('female', 'asthma', 'liver_mild', 'renal', 'pulmonary',
+          'neurological', 'liver_mod_severe', 'malignant_neoplasm')
+
+
+# focus on
+excerpt <- data_
+
+
+# the binary fields must be converted to numeric fields
+binary <- c('asthma', 'liver_mild', 'renal', 'pulmonary', 'neurological', 'liver_mod_severe', 'malignant_neoplasm')
+for (name in binary) {
+  excerpt[, name] <- dplyr::if_else(excerpt[, name] == 'yes', true = 1, false = 0, missing = NULL)
+}
+
+
+# sex
+excerpt$female <- dplyr::if_else(excerpt$sex == 'Female', true = 1, false = 0, missing = NULL)
+
+
+# age group median
+excerpt <- AgeGroupMedians(blob = excerpt)
+
+
+# the time & status
+time <- excerpt$time_to_outcome
+status <- excerpt$deceased
+
+
+# global tau
+times <- function(key) {
+  value <- min(max(excerpt[excerpt[, key] == 1, 'time_to_outcome']),
+      max(excerpt[excerpt[, key] == 0, 'time_to_outcome']))
+  return(value)
+}
+minima <- as.numeric(lapply(X = keys, FUN = function (x){times(x)}))
+tau <- min(minima)
+
+
+# unadjusted
+undaj <- Unadjusted(excerpt = excerpt, time = time, status = status, tau = tau, keys = keys)
+
+
+# adjusted
+adj <- Adjusted(excerpt = excerpt, time = time, status = status, tau = tau, keys = keys)
+
+
+
+
+# diagnostics
+diagnostics <- merge(x = undaj, y = adj, by = 0, all.x = TRUE)
+names(diagnostics)[names(diagnostics) == 'Row.names'] <- 'label'
+
+
[…]

### 2022-01-20 R/background/CorrelationOfPredictors.R
Update R/background/CorrelationOfPredictors.R

@@ -20,15 +20,16 @@ CorrelationOfPredictors <- function (predictors, data) {
   design <- mapply(FUN = IsCorrelatedCC, reference = predictors, MoreArgs = list(variables = predictors, frame = data))
   measures <- dplyr::bind_rows(design)
 
-  # matrix graph
+  # matrix graph, colors = c('#E46726', 'white', 'black'),
   ggcorrplot::ggcorrplot(corr = measures,
                          type = 'upper',
                          ggtheme = ggplot2::theme_minimal(),
-                         colors = c('#E46726', 'white', 'black'),
                          outline.color = 'white',
                          lab = TRUE,
                          lab_size = 3,
                          lab_col = 'grey',
-                         digits = 2)
+                         digits = 2) +
+    scale_fill_gradient2(breaks = c(0, 1), limit = c(0, 1)) +
+    labs(fill = "Cramer's V\n")
 
 }

### 2022-01-20 R/background/Numbers.R
Update R/background/Numbers.R

@@ -8,15 +8,25 @@
 DeceasedNumbers <- function (data) {
 
   deaths <- table(data$deceased, useNA = 'always')
-  return(data.frame(yes = deaths['1'],
-                    no = deaths['0'],
-                    unknown = deaths[length(deaths)],
-                    row.names = 'deceased'))
+  deceased <- data.frame(variable = 'Deceased', yes = deaths['1'], no = deaths['0'],
+                         unknown = deaths[length(deaths)], total = sum(deaths))
+
+  deceased <- deceased %>%
+    mutate(yes_ = yes/total, no_ = no/total, unknown_ = unknown/total)
+
+  deceased <- deceased %>%
+    select(variable, yes, no, unknown, yes_, no_, unknown_)
+
+  return(deceased)
+
+
 }
+
 
 #' Demographics
 DemographicsNumbers <- function (data) {
 
+  # numbers
   demographics <- data %>%
     select(sex, age_group) %>%
     group_by(sex, age_group) %>%
@@ -24,21 +34,29 @@ DemographicsNumbers <- function (data) {
     pivot_wider(id_cols = age_group, names_from = sex, values_from = N) %>%
     data.frame()
 
+  # female ?
   demographics$Total <- select(demographics, !age_group) %>% rowSums()
-  names(demographics) <- c('Age Group', 'Male', 'Female', 'Unknown', 'Total')
+  names(demographics) <- c('variable', 'no', 'yes', 'unknown', 'total')
+
+  # percentages
+  demographics <- demographics %>%
+    mutate(no_ = no/total, yes_ = yes/total, unknown_ = unknown/total)
+  demographics <- demographics %>%
+    select(variable, yes, no, unknown, yes_, no_, unknown_)
 
   return(demographics)
 
 }
+
 
 #' Diseases
 DiseaseNumbers <- function (data) {
 
   disease <- c('asthma', 'liver_mild', 'liver_mod_severe',
             'malignant_neoplasm', 'neurological', 'pulmonary', 'renal')
-  name <- c('Asthma', 'Mild Liver Disease', 'Moderate, Severe Liver Disease',
-            'Malignant Neoplasm', 'Neurological Disorder', 'Pulmonary Disease', 'Renal Disease')
-  labels <- data.frame(disease = disease, Comorbidity = name)
+  name <- c('Asthma', 'Mild Liver\nDisease', 'Moderate, Severe\nLiver Disease',
[…]

### 2022-01-20 R/evaluation/KaplanMeier.R
Update R/evaluation/KaplanMeier.R

@@ -30,7 +30,7 @@ KaplanMeier <- function () {
     data = data_,
     caption = str_glue('B. {nrow(data_)} records'))
 
-  left + right
+  left + patchwork::plot_spacer() + right
 
 }
 

### 2022-01-20 R/restricted/interface.R
Update R/restricted/interface.R

@@ -4,78 +4,8 @@
 # Created on: 06/01/2022
 
 
+source(file = 'R/restricted/RMST.R')
 
-#' Programs
-#'
-source(file = 'R/functions/ExtensiveStudyData.R')
-source(file = 'R/restricted/AgeGroupMedians.R')
-source(file = 'R/restricted/Unadjusted.R')
-source(file = 'R/restricted/Adjusted.R')
+diagnostics <- RMST()
+diagnostics
 
-
-# data
-dataframes <- ExtensiveStudyData(upload = TRUE)
-data_ <- dataframes$data_
-
-
-# for univariate RMST
-keys <- c('female', 'asthma', 'liver_mild', 'renal', 'pulmonary',
-          'neurological', 'liver_mod_severe', 'malignant_neoplasm')
-
-
-# focus on
-excerpt <- data_
-
-
-# the binary fields must be converted to numeric fields
-binary <- c('asthma', 'liver_mild', 'renal', 'pulmonary', 'neurological', 'liver_mod_severe', 'malignant_neoplasm')
-for (name in binary) {
-  excerpt[, name] <- dplyr::if_else(excerpt[, name] == 'yes', true = 1, false = 0, missing = NULL)
-}
-
-
-# sex
-excerpt$female <- dplyr::if_else(excerpt$sex == 'Female', true = 1, false = 0, missing = NULL)
-
-
-# age group median
-excerpt <- AgeGroupMedians(blob = excerpt)
-
-
-# the time & status
-time <- excerpt$time_to_outcome
-status <- excerpt$deceased
-
-
-# global tau
-times <- function(key) {
-  value <- min(max(excerpt[excerpt[, key] == 1, 'time_to_outcome']),
-      max(excerpt[excerpt[, key] == 0, 'time_to_outcome']))
-  return(value)
-}
-minima <- as.numeric(lapply(X = keys, FUN = function (x){times(x)}))
-tau <- min(minima)
-
-
-# unadjusted
-undaj <- Unadjusted(excerpt = excerpt, time = time, status = status, tau = tau, keys = keys)
-
-
-# adjusted
-adj <- Adjusted(excerpt = excerpt, time = time, status = status, tau = tau, keys = keys)
-
-
-
-
-# diagnostics
-diagnostics <- merge(x = undaj, y = adj, by = 0, all.x = TRUE)
-names(diagnostics)[names(diagnostics) == 'Row.names'] <- 'label'
-
-
[…]

### 2022-01-20 R/main.R
Update R/main.R

@@ -8,6 +8,7 @@
 #' Programs
 #'
 source(file = 'R/functions/ExtensiveStudyData.R')
+source(file = 'R/background/Numbers.R')
 source(file = 'R/demographics/Is.R')
 
 
@@ -35,3 +36,5 @@ table(data$age_group)
 #' of England's population demographics
 #'
 Is()
+
+

### 2022-01-20 R/restricted/RMST.R
Add R/restricted/RMST.R

@@ -1,0 +1,89 @@
+# Title     : RMST.R
+# Objective : Restricted Mean Survival Time
+# Created by: greyhypotheses
+# Created on: 19/01/2022
+
+
+source(file = 'R/functions/ExtensiveStudyData.R')
+source(file = 'R/restricted/AgeGroupMedians.R')
+source(file = 'R/restricted/Unadjusted.R')
+source(file = 'R/restricted/Adjusted.R')
+
+
+RMST <- function () {
+
+
+  # data
+  dataframes <- ExtensiveStudyData(upload = TRUE)
+  data_ <- dataframes$data_
+
+
+  # for univariate RMST
+  keys <- c('female', 'asthma', 'liver_mild', 'renal', 'pulmonary',
+            'neurological', 'liver_mod_severe', 'malignant_neoplasm')
+
+
+  # focus on
+  excerpt <- data_
+
+
+  # the binary fields must be converted to numeric fields
+  binary <- c('asthma', 'liver_mild', 'renal', 'pulmonary', 'neurological', 'liver_mod_severe', 'malignant_neoplasm')
+  for (name in binary) {
+    excerpt[, name] <- dplyr::if_else(excerpt[, name] == 'yes', true = 1, false = 0, missing = NULL)
+  }
+
+
+  # sex
+  excerpt$female <- dplyr::if_else(excerpt$sex == 'Female', true = 1, false = 0, missing = NULL)
+
+
+  # age group median
+  excerpt <- AgeGroupMedians(blob = excerpt)
+
+
+  # the time & status
+  time <- excerpt$time_to_outcome
+  status <- excerpt$deceased
+
+
+  # global tau
+  times <- function(key) {
+    value <- min(max(excerpt[excerpt[, key] == 1, 'time_to_outcome']),
+                 max(excerpt[excerpt[, key] == 0, 'time_to_outcome']))
+    return(value)
+  }
+  minima <- as.numeric(lapply(X = keys, FUN = function (x){times(x)}))
+  tau <- min(minima)
+
+
+  # unadjusted
+  undaj <- Unadjusted(excerpt = excerpt, time = time, status = status, tau = tau, keys = keys)
+
+
+  # adjusted
+  adj <- Adjusted(excerpt = excerpt, time = time, status = status, tau = tau, keys = keys)
+
+
+  # diagnostics
+  diagnostics <- merge(x = undaj, y = adj, by = 0, all.x = TRUE)
[…]

### 2023-10-21 R/background/Numbers.R
Update R/background/Numbers.R

@@ -54,8 +54,8 @@ DiseaseNumbers <- function (data) {
 
   disease <- c('asthma', 'liver_mild', 'liver_mod_severe',
             'malignant_neoplasm', 'neurological', 'pulmonary', 'renal')
-  name <- c('Asthma', 'Mild Liver\nDisease', 'Moderate, Severe\nLiver Disease',
-            'Malignant Neoplasm', 'Neurological\nDisorder', 'Pulmonary Disease', 'Renal Disease')
+  name <- c('Asthma', 'Mild Liver Disease', 'Moderate, Severe Liver Disease',
+            'Malignant Neoplasm', 'Neurological Disorder', 'Pulmonary Disease', 'Renal Disease')
   labels <- data.frame(disease = disease, variable = name)
 
   diseases <- data %>%

### 2023-10-21 manuscript.Rmd
Update manuscript.Rmd

--- 2022-01-18 manuscript.Rmd
@@
 knitr::opts_chunk$set(echo = FALSE)
-options(knitr.table.format = "latex")
+options(knitr.table.format = 'latex')
+options(knitr.kable.NA = '')
 ```
@@
 library(stringr)
+library(matrixStats)
 
@@
 library(survival)
+library(survRM2)
 library(patchwork)
+library(kableExtra)
 ```
@@
 ```{r}
-sys.source(file = 'R/evaluation/KaplanMeier.R', envir = knitr::knit_global())
-sys.source(file = 'R/functions/StudyData.R', envir = knitr::knit_global())
+sys.source(file = 'R/functions/ExtensiveStudyData.R', envir = knitr::knit_global())
 
 sys.source(file = 'R/background/Numbers.R', envir = knitr::knit_global())
+sys.source(file = 'R/background/CorrelationOfPredictors.R', envir = knitr::knit_global())
+
+sys.source(file = 'R/missing/Pattern.R', envir = knitr::knit_global())
+sys.source(file = 'R/missing/Test.R', envir = knitr::knit_global())
+
+sys.source(file = 'R/demographics/Is.R', envir = knitr::knit_global())
+
+sys.source(file = 'R/modelling/UnivariateAnalysis.R', envir = knitr::knit_global())
+sys.source(file = 'R/modelling/MultivariateAnalysis.R', envir = knitr::knit_global())
+
+sys.source(file = 'R/evaluation/KaplanMeier.R', envir = knitr::knit_global())
+
+sys.source(file = 'R/restricted/RMST.R', envir = knitr::knit_global())
 ```
@@
 >
-> **Design** *Prospective cohort study; ongoing.*
+> **Design & Setting** *The underlying study is the International Severe Acute Respiratory and Emerging Infections Consortium (ISARIC) World Health Organization (WHO) Clinical Characterisation Protocol UK (CCP-UK) study, which enrols in-patients at at-least 260 England, Scotland, and Wales hospitals; the ISARIC Coronavirus Clinical Characterisation Consortium (ISARIC 4C) conducts the study.
+It is an ongoing prospective cohort study.
+Herein, the patients are patients enrolled in a participating England hospital between 10 February 2020 and 5 July 2020.*
 >
-> **Setting** *The International Severe Acute Respiratory and Emerging Infections Consortium (ISARIC) World Health Organization (WHO) Clinical Characterisation Protocol UK (CCP-UK) study, which enrols in-patients at at-least 260 England, Scotland, and Wales hospitals; the ISARIC Coronavirus Clinical Characterisation Consortium (ISARIC 4C) conducts the study.
-The investigation's patients are patients enrolled in a participating England hospital between 10 February 2020 and 5 July 2020.*
+> **Results** The results of the investigation are inconclusive.
+Modelling and analysis could not proceed via the Cox Proportional Hazards model due to the violation of the proportionality assumption.
+The restricted mean survival time analysis suggest that sex, malignant neoplasm, neurological disorder, pulmonary disease, and renal disease might have a significant impact on in-hospital coronavirus 19 disease survival.
 >
-> **Participants** *Each study enrollee has a confirmed SARS-CoV-2 infection, or a high infection likelihood.  The investigation's patients are members of age groups *30 - 39*, *40 - 49*, and higher.*
->
-> **Outcome** Patient survival.
->
-> **Results** Inconclusive.
->
-> **Conclusions**  Insufficient spectrum of factors.
+> **Conclusions**  The investigation should be repeated with much more appropriate data, e.g., continuous age variables, start & stop times of events, follow-ups, etc.
+The analysis of missing values hints at flaws in data collection, this might be partly addressed via a clearer recruitment protocol.
 
-# Introduction
+\vspace{35pt}
 
-During the past 2 years, infections due to severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2) have led to a high number of coronavirus 19 disease deaths and in-hospital care.
-With time, case fatality rates, and in-hospital survival rates, have improved [...] as our understanding of coronavirus 19 disease pathogenesis has developed.
-Consequently,  disease prevention, mitigation, and treatment measures have improved.
+# INTRODUCTION
 
-Effective disease treatment measures depend on our understanding of the disease's prognosis factors [@riley2013], the focus of prognostic factors research.
+During the past 2 years, infections due to severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2) have led to a high number of coronavirus 19 disease deaths and in-hospital care [@jhu2020, @ritchie2020].
+With time, in-hospital survival rates, have improved [@docherty2021].
+A number of factors underlie the improvements, one of the most crucial being a better understanding of coronavirus 19 disease, which has led to much more effective treatment measures [@docherty2021].
+
+Effective disease treatment measures partly depend on our understanding of the disease's prognosis factors [@riley2013], the focus of prognostic factors research.
 Prognostic factor research *aims to identify factors associated with subsequent clinical outcome in people with a particular disease or health condition* [@riley2013].
 
-This brief investigates the prognostic factor potential of a set of measures associated with a prospective cohort study.
-The study's hypothesis is that *one or more of the measures has a significant impact on each patient's survival probability*.
-The ISARIC 4C study outcomes are: discharged alive, remains in hospital, death, transferred, and palliative discharge.
-Herein, the overarching outcomes of interest are: dead, alive.
-The predictors, factors, under investigation
+The definitive set of factors that affect in-hospital survival probabilities of coronavirus 19 disease patients is unknown, and is a research focus [@gorog2022, @mcgurnaghan2021].
+This brief investigation investigates the prognostic factor potential of a set of measures associated with the ISARIC prospective cohort study.
+The hypothesis being *one or more of the measures has a significant impact on each patient's survival probability*.
 
-# Methods
+\vspace{35pt}
+
+# METHODS
+
+\vspace{10pt}
 
@@
 
-This investigation's data is an indirect ISARIC (International Severe Acute Respiratory Infection Consortium)  4C (Coronavirus Clinical Characterisation Consortium) study data set.
-The study is an ongoing prospective cohort study based in more than 200 United Kingdom acute care hospitals.
-This investigation's data is an England only data set.
-The reporting ...
+This investigation's data is an indirect ISARIC (International Severe Acute Respiratory Infection Consortium) 4C (Coronavirus Clinical Characterisation Consortium) study data set; the data set is a project assigned data set, not actively collected.
+The ISARIC 4C study is an ongoing prospective cohort study that is conducted by the ISARIC Coronavirus Clinical Characterisation Consortium (ISARIC 4C).
+The study enrols in-patients at at-least 260 England, Scotland, and Wales acute care hospitals.
+Herein, the patients are patients enrolled in a participating England hospital between 10 February 2020 and 5 July 2020, i.e., this paper's investigation focuses on an England only data set.
+Additionally, the England only data set consists of patients within age groups *30 - 39*, *40 - 49*, and higher.*
 
-\vspace{20pt}
-
-## Participants
-
-Inclusion criteria were people of all ages who were admitted to one of 208 acute care hospitals in England, Scotland, and Wales with proven or high likelihood of infection with a pathogen of public health interest, defined as SARS-CoV-2 for this event by Public Health England.
-Reverse transcriptase polymerase chain reaction was the only mode of testing available during the period of study.
-The decision to test was at the discretion of the clinician attending the patient, and not defined by protocol.
-The enrolment criterion “high likelihood of infection” reflects that a preparedness protocol cannot assume that a diagnostic test will be available for an emergent pathogen.
-Site training emphasises that only patients who tested positive for covid-19 were eligible for enrolment.
-
-National guidance was provided by Public Health England and other UK public health agencies that advised who to test based on clinical case definitions for possible covid-19.
-We also included patients who had been admitted for a separate condition but had tested positive for covid-19 during their hospital stay.
-Patients were only enrolled during their index admission.
-Patients had clinical information from their routine health records uploaded into the case report form.
-Consent was not required for collection of depersonalised routine healthcare data for research in England and Wales.
-A waiver for consent was given by the Public Benefit and Privacy Panel in Scotland.
-
-\vspace{20pt}
-
-## Data
+According to the project brief, each study enrollee has a confirmed SARS-CoV-2 infection, or a high infection likelihood.
+The study collects its baseline data via paper case report forms developed by ISARIC, and the WHO (World Health Organization), for outbreak investigations.
 
@@
 
-The ISARIC 4C study collects its baseline data via paper case report forms developed by ISARIC, and the WHO (World Health Organization), for outbreak investigations.
-Additionally, the data brief states that the baseline data records are *uploaded [on] admission, and usually before hospital episodes were complete, to a REDCap database (Research Electronic Data Capture, Vanderbilt University, US, hosted by University of Oxford, UK).*
-
 \vspace{20pt}
 
-## Exploratory Data Analysis
+## Outcomes
 
-* Relationships
-* Correlations
+The ISARIC 4C study outcomes are: discharged alive, remains in hospital, death, transferred, and palliative discharge.
+Herein, the overarching outcomes of interest are: dead, alive.
+Hence, investigation's censoring logic is
+
+> Right Censored:
+>
+>  * discharged alive
+>
+>  * remains in hospital
+>
+>  * transferred
+>
+>  * palliative discharge
+
+> Uncensored:
+>
+>  * death
+
+Without more information it will be ill-advised to assume that death is the final outcome of patients released under palliative discharge.
+Especially because in addition to *end of life care*, palliative care also includes *supportive care* for curable diseases, i.e., support for patients that are expected to recover, but who need critical support whilst convalescing.
+[@apm2019]
 
@@
 
-* missing completely at random (MCAR) / administrative errors, accidents
-* missing at random (MAR) / missing data associated with known patient characteristics [independent variables], or the outcome
-* missing not at random (MNAR) / missing data associated with missing values of the factor/predictor in question or with unobserved predictors
+\vspace{10pt}
+
+* Missing Completely at Random (MCAR): administrative errors, accident.
+
+* Missing at Random (MAR): missing data associated with known patient characteristics [independent variables], or the outcome.
+
+* Missing Not at Random (MNAR): missing data associated with missing values of the factor/predictor in question or with unobserved predictors.
+
+\vspace{10pt}
 
@@
 
-Herein, missing data mechanisms analysis is via logistic regression.
-Almost all the independent variables have missing values patterns that violate MCAR; addressed via multiple imputation in-line with the advice and protocols of [@sterne2009] & [@steyerberg2010].
-
-## Factor Analysis
-
-The Cox Hazard, Restricted Mean Survival
+The missing data mechanisms analysis [Appendix] suggests that almost all the independent variables have missing values patterns that violate MCAR.
+Consequently, a multiple imputation step, in-line with the advice and protocols of [@sterne2009] & [@steyerberg2010], is part of this investigation.
 
@@
 
-# Results
+## Modelling & Analysis: Potential Prognostic Factors
 
-## Characteristics
+The potential, candidate, prognostic factors are sex, age group, asthma, mild liver disease, moderate/severe liver disease, renal disease, chronic pulmonary disease, chronic neurological disorder, and malignant neoplasm.
+Their effects, or otherwise, on survival probabilities, are explored via the Cox Proportional Hazard Model [@cox1972] and the Restricted Mean Survival Time [@royston2013].
 
-\begin{table}[h!]
-    \centering
-    \renewcommand{\arraystretch}{1.25}
-    \begin{tabular}{l r r r}
-    & & & \\
-    \textbf{In Hospital Mortality} & & & \\
-    \hline
-     & \textbf{no} & \textbf{yes} & \textbf{NA} \\
-    deceased & 31724 & 14049 & 347 \\
-     & 0.688 & 0.305 & 0.008 \\
-     & & & \\
-    \textbf{Comorbidities} & & & \\
-    \hline
-     & \textbf{no} & \textbf{yes} & \textbf{NA} \\
-    asthma & 37376 & 5872 & 2872 \\
-    & 0.81 & 0.13 & 0.06 \\
-    mild liver disease & 42021 & 666 & 3433 \\
-     & 0.91 & 0.01 & 0.07 \\
-    renal disease & 35158 & 8056 & 2906 \\
-     & 0.76 & 0.17 & 0.06 \\
-    pulmonary & 35265 & 8078 & 2777 \\
-     & 0.76 & 0.18 & 0.06 \\
-    neurological disorder & 37227 & 5806 & 3087 \\
-     & 0.81 & 0.13 & 0.07 \\
-    liver mod severe &  41901 &  902 & 3317 \\
-     & 0.91 & 0.02 & 0.07 \\
-    malignant neoplasm & 38164 & 4734 & 3222 \\
-     & 0.83 & 0.10 & 0.07 \\
-     & & & \\
-    \textbf{Demographics} & & & \\
-    \hline
-    \textbf{Age Group} &  \textbf{Male} & \textbf{Female} & \textbf{NA} \\
-    80-89 & 7067 & 6197 & 31 \\
-    & 0.153 & 0.134 & 0.001 \\
-    70-79 & 6388 & 4445 & 19 \\
-    & 0.139 & 0.096 & 0.000 \\
-    60-69 & 4321 & 2601 & 9 \\
-    & 0.094 & 0.056 & 0.000 \\
-    50-59 & 3463 & 2151 & 9 \\
-    & 0.075 & 0.047 & 0.000 \\
-    90+ & 2085 & 2816 & 13 \\
-    & 0.045 & 0.061 & 0.000 \\
-    40-49 & 1846 & 1110 & 5 \\
-    & 0.040 & 0.024 & 0.000 \\
-    30-39 & 784 & 756 & 4 \\
-    & 0.017 & 0.016 & 0.000 \\
-    \hline
-    \end{tabular}
-\end{table}
+\pagebreak
+
+# RESULTS
+
+\vspace{10pt}
+
+## Descriptive Statistics
+
+The numbers of *Table 1* summarise the death, comorbidity , and demographic numbers of the investigation's data set.
+Of the received data set of $50,000$ records, $3,880$ are considered implausible because each has an outcome date that precedes its admission date.
+Consequently, each outlined modelling and analysis step is due to $46,120$ records.
+
+\footnotesize
+```{r}
+special <- SpecialNumbers(data = data)
+
+special %>%
+  dplyr::mutate_all(linebreak) %>%
+  kableExtra::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    digits = 2,
+    escape = FALSE,
+    row.names = FALSE,
+    col.names = kableExtra::linebreak(c('', 'yes', 'no', 'unknown', 'yes', 'no', 'unknown' )),
+    caption = 'Data Description')  %>%
+  kableExtra::kable_styling(position = 'center') %>%
+  kableExtra::add_header_above(c('', 'N' = 3, 'Percentage' = 3)) %>%
+  kableExtra::column_spec(column = 1, latex_column_spec = "p{0.17\\\\linewidth}l")
+
+```
+\normalsize
+
+\pagebreak
+
+The graph of *fig.
+1* illustrates the degree of correlation between the potential prognostic factors.
+The embedded Cramer's V values suggest weak associations only; the degree of freedom per cell is one.
+
+\vspace{20pt}
+
+```{r out.width = '65%', fig.align = 'center', results = 'hide', warning = FALSE, message = FALSE, fig.cap = 'Is there a degree of correlation between predictors?'}
+CorrelationOfPredictors(
+  predictors = c('age_group', 'sex', 'asthma', 'liver_mild', 'renal', 'pulmonary',
+                 'neurological', 'liver_mod_severe', 'malignant_neoplasm'),
+  data = data)
+```
+
+\vspace{20pt}
 
@@
 
-```{r fig.height = 3.5, fig.align = 'center', warning = FALSE, message = FALSE, fig.cap = "The null survival curves, whereby (A) is based on complete cases only, whilst (B) is based on all plausible records after multiple imputation."}
+The graphs of *fig 2* illustrate the survival proportions over time.
+The probability of survival drops sharply during the first 200 days, this unsteady pattern suggest that the risk of death over time might not be constant.
+
+```{r fig.height = 3.5, out.width = '85%', fig.align = 'center', warning = FALSE, message = FALSE, fig.cap = "The null survival curves, whereby (A) is based on complete cases only, whilst (B) is based on all plausible records after multiple imputation."}
 KaplanMeier()
@@
 
-## Null Survival Curve
+\pagebreak
 
-* Complete Case
-* Imputation Case
+## Univariate Analysis: Cox Proportional Hazards Model
 
-## Uni-variate & Multivariate Analysis
+This Cox Proportional Hazard Model uni-variate analysis models & analyses the separate effect of each potential prognostic factor on survival probability.
+The analysis is via the imputed data set, which excludes the *3,880* implausible observation, i.e., there are *46,120* records; *50,000* - *3,880*.
 
-* The table ... Cox
+In terms of age groups, the reference age group is *80 – 89*, and their $p_{value}$ scores suggests that **(a)** each age group, except *90+*, has a lower risk of death compared to age group *80 – 89*, and **(b)** the risk of death increases with age.
+The sex $p_{value}$ suggests that females have a lower risk of death compared to males.
+The $p_{value}$ scores for patients with an asthma, renal disease, pulmonary disease, or malignant neoplasm, diagnosis suggests that these patients have a higher risk of death compared to their respective disease/disorder free counterparts.
 
-## Adjusted Analysis via Restricted Mean Survival Time
+Alas, the proportionality tests invalidate most of the preceding observations; age group, malignant neoplasm, and neurological disorder have a score test value below $0.05$, i.e., the proportionality assumption does not hold.
+The multivariate analysis confirms this observation.
 
-The proportional hazard assumption does not hold ...
+\vspace{20pt}
 
-# Discussion
+\footnotesize
+```{r}
+diagnostics <- UnivariateAnalysisData()
+
+diagnostics %>%
+  dplyr::mutate_all(linebreak) %>%
+  kableExtra::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    digits = 3,
+    escape = FALSE,
+    col.names = kableExtra::linebreak(c('', 'coefficient', 'hazard\nratio', 'lower',
+                                        'upper', '$p_{value}$', '${U}$' )),
+    caption = 'Univariate Analysis: Imputed Data', position = '!ht') %>%
+  kableExtra::kable_styling(position = 'center') %>%
+  kableExtra::add_header_above(c('', '', '', 'hazard ratio\nCI' = 2, '', '')) %>%
+  kableExtra::add_footnote(c('CI: Confidence Interval',
+                             'U: Proportionality score test'), notation = "number")  %>%
+  kableExtra::column_spec(column = 1, latex_column_spec = "p{0.20\\\\linewidth}l") %>%
+  kableExtra::column_spec(column = 2, latex_column_spec = "r")
+```
+\normalsize
+
+\pagebreak
+
+## Multivariate Analysis: Cox Proportional Hazards Model
+
+Similar to the univariate analysis, the proportionality tests invalidate the multivariate analysis results of *table 3*; age group, malignant neoplasm, and neurological disorder have score test values below $0.05$.
+
+\footnotesize
+```{r}
+diagnostics <- MultivariateAnalysisData()
+
+diagnostics %>%
+  dplyr::mutate_all(linebreak) %>%
+  kableExtra::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    digits = 3,
+    escape = FALSE,
+    col.names = kableExtra::linebreak(c('', 'coefficient', 'hazard\nratio', 'lower',
+                                        'upper', '$p_{value}$', '${U}$' )),
+    caption = 'Multivariate Analysis: Imputed Data', position = '!ht') %>%
+  kableExtra::kable_styling(position = 'center') %>%
+  kableExtra::add_header_above(c('', '', '', 'hazard ratio\nCI' = 2, '', '')) %>%
+  kableExtra::add_footnote(c('CI: Confidence Interval',
+                             'U: Proportionality score test'), notation = "number")  %>%
+  kableExtra::column_spec(column = 1, latex_column_spec = "p{0.20\\\\linewidth}l") %>%
+  kableExtra::column_spec(column = 2, latex_column_spec = "r")
+```
+\normalsize
+
+\pagebreak
+
+## Unadjusted & Adjusted Analysis: Restricted Mean Survival Time
+
+The restricted mean survival time (RMST) is an alternative survival analysis approach, especially when the risk over time might be inconstant and/or the Cox Proportional Hazards Model assumptions do not hold.
+The restricted mean survival time is *...
+a measure of average survival from time 0 to a specified time point ...* [@royston2013]  The results of *Table ...* summarise the unadjusted and adjusted uni-variate RMST analysis of the named predictors.
+
+In the case of the adjusted analysis, a continuous variable field of age group medians, instead of the age group field, is used; access to the original age values would have been much better.
+As usual, the adjusted analysis is the case whereby a single predictor is in focus whilst controlling for other variables/predictors, and the hypotheses per predictor are
+
+\vspace{15pt}
+
+* $H_{0}: \tau_{yes} = \tau_{no}$
+* $H_{1}: \tau_{yes} \neq \tau_{no}$
+
+\vspace{15pt}
+
+Wherein the null hypothesis is that the survival time w.r.t.
+(with respect to) yes responses is equivalent to the survival time w.r.t.
+no responses.
+Noting that the specified time point is 267 days $\longrightarrow$ minimum of the *maximum time to outcome* across all predictors
+
+\vspace{15pt}
+
+```r
+  # for a single predictor
+  min(max(data[data[, predictor] == 1, 'time_to_outcome']),
+      max(data[data[, predictor] == 0, 'time_to_outcome']))
+```
+
+\vspace{15pt}
+
+Then the **adjusted analysis results** of *table 4* suggests that within a 267 days limit,
+
+* on average female patients survived 18.46 days longer than males.
+* the survival days of patients with malignant neoplasm is approximately 18 days shorter than those without.
+* on average patients with a neurological disorder had approximately 31 fewer survival days.
+* on average patients with a pulmonary disease had approximately 12 fewer survival days.
+* on average patients with a renal had approximately 13 fewer survival days.
+
+There are no other significant results.
+Further analysis, much more appropriate data, e.g., continuous age values, and a wider range of potential factors [@docherty2021], is required in order to make definitive conclusions.
+
+\footnotesize
+```{r}
+diagnostics <- RMST()
+
+diagnostics %>%
+  dplyr::mutate_all(linebreak) %>%
+  kableExtra::kable(
+    format = 'latex',
+    booktabs = TRUE,
+    digits = 2,
+    escape = FALSE,
+    col.names = kableExtra::linebreak(c('', 'est.', 'lower', 'upper', '$p_{value}$',
+                                        'est.', 'lower', 'upper', '$p_{value}$')),
+    caption = 'Restricted Mean Survival Time Modelling and Analysis', position = '!ht') %>%
+    kableExtra::kable_styling(position = 'center') %>%
+    kableExtra::add_header_above(c('', '', 'CI' = 2, '', '', 'CI' = 2, '')) %>%
+    kableExtra::add_header_above(c('', 'Unadjusted' = 4, 'Adjusted' = 4)) %>%
+    kableExtra::column_spec(column = 1, latex_column_spec = "p{0.20\\\\linewidth}l")
+```
+\normalsize
+
+\vspace{65pt}
+
+\pagebreak
+
+# DISCUSSION
 
@@
 
-**Proportional Hazards**: The curves.
+**Proportional Hazards**: The proportionality tests illustrated that the data violates the proportionality assumption.
+
+\vspace{20pt}
 
@@
 
-The investigation opted for the latter, especially because ...
+The investigation opted for the latter, due to the missing data patterns observed (Appendix)
+
+\vspace{20pt}
 
@@
 
-Possible bias points
+There are a few possible bias points.
+By virtue of the study's design pre-study--admission differences between hospitals is possible due to the ambiguity of the study's patient recruitment terms, and this might lead to post-study differences.
+In a nutshell, the study is susceptible to selection bias.
+[Trochim]  Bias due to missing data [@sterne2009].
 
-* By virtue of the study's design pre-study--admission differences between hospitals is possible due to the ambiguity of the study's patient recruitment terms, and this might lead to post-study differences.
-In a nutshell, the study is susceptible to selection bias.
-[Trochim]
+\vspace{20pt}
 
-* Missing data [@sterne2009]
+## NHS England COVID-19 Hospital Activity & ISARIC Study Excerpt
 
-# Conclusion
+It is probably impossible to compare the study's patient population to the NHS England patient population of the same
+time period.  There are a few reasons why $\longrightarrow$ the COVID-19 Hospital Activity Data:
 
+* Does not include a daily record of new COVID-19 patients admissions.  Instead, it continuously updates a number
+that denotes *the number of admitted COVID-19 patients that were still in hospital during the last 24 hours*
+* Does not include in-hospital COVID-19 deaths.
+* Has a different age groups demarcation: `[0  5]`, `[6  17]`, `[18  64]`, `[65  84]`, `85+`
+
+Hence, even if the results had been conclusive, inference to the population might have been tricky.
+
+\pagebreak
+
+# CONCLUSION
+
+The results of the investigation are inconclusive.
+Modelling and analysis could not proceed via the Cox Proportional Hazards model due to the violation of the proportionality assumption.
+The restricted mean survival time analysis suggest that sex, malignant neoplasm, neurological disorder, pulmonary disease, and renal disease might have a significant impact on in-hospital coronavirus 19 disease survival.
 The analysis of missing values hints at flaws in data collection.
+
+\pagebreak
 
@@
 
-```{r}
+The study's female/male population is not representative of England's population, therefore population inference is not appropriate.
 
+```{r fig.height = 3.5, out.width = '49%', fig.align = 'center', warning = FALSE, message = FALSE, fig.cap = "Is the investigation's population representative of England's population?"}
+Is()
 ```
 
-Graphs
+\vspace{10pt}
 
@@
 
-Clusters, patterns, associations.
+The clusters, patterns, associations, in relation to missing values.
 
+\vspace{10pt}
+
+```{r out.width = '65%', fig.align = 'center', fig.cap = 'Fractions denoting the instance proportions whereby a pair of predictors are both NA values'}
+PatternDendogramNA(data = data)
+```
+
+\vspace{10pt}
+
+```{r out.width = '65%', fig.align = 'center', fig.cap = 'The proportion of missing values per predictor variable'}
+PatternVariableNA(data = data)
+```
+
+\vspace{10pt}
+
+```{r warning = FALSE, message = FALSE, include = FALSE}
+diagnostics <- Test(data = data)
+
+diagnostics %>%
+  dplyr::mutate_all(linebreak) %>%
+    kableExtra::kable(
+      format = 'latex',
+      booktabs = TRUE,
+      digits = 2,
+      escape = FALSE,
+      row.names = FALSE,
+      col.names = kableExtra::linebreak(c('Missing Value\nCells of', 'Associated with one or more\naux./pred. variables' )),
+      caption = 'Data Description') %>%
+    kableExtra::column_spec(column = 1, latex_column_spec = "p{0.3\\\\linewidth}l")
+```
+
+\pagebreak
+\pagebreak
 <!--- # References {#references .unnumbered} --->

### 2023-10-21 R/demographics/Is.R
Update R/demographics/Is.R

@@ -42,6 +42,7 @@ Is <- function () {
     ggplot() +
     geom_bar(mapping = aes(x = age_group, y = `Female/Male Ratio`, fill = Case),
              stat = 'identity', position = position_dodge2(), alpha = 0.35) +
+    scale_fill_manual(values = c('orange', 'black')) +
     theme_minimal() +
     theme(panel.grid.minor = element_blank(),
           panel.grid.major = element_line(size = 0.15),

### 2023-10-21 R/evaluation/KaplanMeier.R
Update R/evaluation/KaplanMeier.R

@@ -30,7 +30,7 @@ KaplanMeier <- function () {
     data = data_,
     caption = str_glue('B. {nrow(data_)} records'))
 
-  left + patchwork::plot_spacer() + right
+  left + right
 
 }
 

### 2023-10-21 R/missing/Imputation.R
Update R/missing/Imputation.R

@@ -9,6 +9,10 @@
 #'
 #' @param training: the training data set
 #'
+#' @note https://www.rdocumentation.org/packages/mice/versions/3.15.0/topics/mice
+#'       https://www.gerkovink.com/miceVignettes/Passive_Post_processing/Passive_imputation_post_processing.html
+#'       Try a named method list, especially for explicit exlusions of columns
+#'
 Imputation <- function (data) {
   imputation <- mice::mice(data = data, m = 13, maxit = 10, seed = 5)
   return(imputation)

### 2023-10-21 R/missing/interface.R
Update R/missing/interface.R

@@ -44,3 +44,8 @@ instances <- data[, variables]
 details <- mapply(FUN = MechanismTest, predictor = focus,
                   MoreArgs = list(instances = instances))
 data.frame(t(details))
+
+calculations <- data.frame(t(details)) %>%
+  mutate(na_of = row.names(.))
+calculations %>%
+  select(na_of, associations, elements)

### 2023-10-21 R/modelling/MultivariateAnalysis.R
Update R/modelling/MultivariateAnalysis.R

@@ -43,6 +43,22 @@ MultivariateAnalysisData <- function () {
 
   analysis <- Multivariate(blob = data_)
   diagnostics <- analysis$diagnostics
+  diagnostics <- diagnostics %>% arrange(label)
+
+  # names, labels, indices
+  settings <- data.frame(label = c('age_group70-79', 'age_group60-69', 'age_group50-59', 'age_group90+', 'age_group40-49',
+                                   'age_group30-39', 'sexFemale', 'asthmayes', 'liver_mildyes', 'renalyes', 'pulmonaryyes',
+                                   'neurologicalyes', 'liver_mod_severeyes', 'malignant_neoplasmyes'),
+                         name = c('Age Group 70 - 79', 'Age Group 60 - 69', 'Age Group 50 - 59', 'Age Group 90+',
+                                  'Age Group 40 - 49', 'Age Group 30 - 39', 'Sex Female', 'Asthma Yes',
+                                  'Liver Disease, (Mild) Yes', 'Renal Disease Yes', 'Pulmonary Disease Yes',
+                                  'Neurological Disorder Yes', 'Liver Disease, (Moderate, severe) Yes', 'Malignant Neoplasm Yes'))
+
+  diagnostics <- dplyr::left_join(x = diagnostics, y = settings, by = 'label')
+  diagnostics <- diagnostics %>%
+    select(name, Coefficient, HR, HRLCI, HRUCI, p_value, proportionality)
+
+  return(diagnostics)
 
 }
 
@@ -54,4 +70,21 @@ MultivariateAnalysisCC <- function () {
 
   C <- data[complete.cases(data),]
   analysis <- Multivariate(blob = C)
+  diagnostics <- analysis$diagnostics
+  diagnostics <- diagnostics %>% arrange(label)
+
+  # names, labels, indices
+  settings <- data.frame(label = c('age_group70-79', 'age_group60-69', 'age_group50-59', 'age_group90+', 'age_group40-49',
+                                   'age_group30-39', 'sexFemale', 'asthmayes', 'liver_mildyes', 'renalyes', 'pulmonaryyes',
+                                   'neurologicalyes', 'liver_mod_severeyes', 'malignant_neoplasmyes'),
+                         name = c('Age Group 70 - 79', 'Age Group 60 - 69', 'Age Group 50 - 59', 'Age Group 90+',
[…]

### 2023-10-21 R/modelling/UnivariateAnalysis.R
Update R/modelling/UnivariateAnalysis.R

@@ -46,6 +46,25 @@ UnivariateAnalysisData <- function () {
   analysis <- Univariate(blob = data_)
   diagnostics <- analysis$diagnostics
 
+  # names, labels, indices
+  diagnostics$label <- row.names(diagnostics)
+  row.names(diagnostics) <- NULL
+  diagnostics <- diagnostics %>% arrange(label)
+
+  settings <- data.frame(label = c('age_group70-79', 'age_group60-69', 'age_group50-59', 'age_group90+', 'age_group40-49',
+                                   'age_group30-39', 'sexFemale', 'asthmayes', 'liver_mildyes', 'renalyes', 'pulmonaryyes',
+                                   'neurologicalyes', 'liver_mod_severeyes', 'malignant_neoplasmyes'),
+                         name = c('Age Group 70 - 79', 'Age Group 60 - 69', 'Age Group 50 - 59', 'Age Group 90+',
+                                  'Age Group 40 - 49', 'Age Group 30 - 39', 'Sex Female', 'Asthma Yes',
+                                  'Liver Disease, (Mild) Yes', 'Renal Disease Yes', 'Pulmonary Disease Yes',
+                                  'Neurological Disorder Yes', 'Liver Disease, (Moderate, severe) Yes', 'Malignant Neoplasm Yes'))
+
+  diagnostics <- dplyr::left_join(x = diagnostics, y = settings, by = 'label')
+  diagnostics <- diagnostics %>%
+    select(name, Coefficient, HR, HRLCI, HRUCI, p_value, proportionality)
+
+  return(diagnostics)
+
 
 
 }
@@ -58,6 +77,26 @@ UnivariateAnalysisCC <- function () {
 
   C <- data[complete.cases(data),]
   analysis <- Univariate(blob = C)
+  diagnostics <- analysis$diagnostics
+
+  # names, labels, indices
+  diagnostics$label <- row.names(diagnostics)
+  row.names(diagnostics) <- NULL
+  diagnostics <- diagnostics %>% arrange(label)
+
+  settings <- data.frame(label = c('age_group70-79', 'age_group60-69', 'age_group50-59', 'age_group90+', 'age_group40-49',
+                                   'age_group30-39', 'sexFemale', 'asthmayes', 'liver_mildyes', 'renalyes', 'pulmonaryyes',
+                                   'neurologicalyes', 'liver_mod_severeyes', 'malignant_neoplasmyes'),
[…]

### 2023-10-21 R/restricted/RMST.R
Update R/restricted/RMST.R

@@ -75,7 +75,7 @@ RMST <- function () {
                                    'neurological', 'liver_mod_severe', 'malignant_neoplasm'),
                          name = c('Female', 'Asthma',
                                   'Liver Disease (Mild)', 'Renal Disease', 'Pulmonary Disease',
-                                  'Neurological Disorder', 'Liver Disease (Moderate, Severe)', 'Malignant Neoplasm Y'))
+                                  'Neurological Disorder', 'Liver Disease (Moderate, Severe)', 'Malignant Neoplasm'))
 
 
   # restricted mean survival calculations

### 2023-10-21 R/testing/interface.R
Add R/testing/interface.R

@@ -1,0 +1,21 @@
+# Title     : interface.R
+# Objective : Interface
+# Created by: greyhypotheses
+# Created on: 09/01/2022
+
+
+#' testing
+#' multiple imputation & complete case
+#'
+
+# Unboosted: Narrow external validation
+predict(object = model, newdata = testing_)
+# calibration plot
+# AUC
+# sensitivity analysis
+
+# Boosted: Narrow external validation
+# mboost::survFit(object = model, newdata = testing_)
+# calibration plot
+# AUC
+# sensitivity analysis
